---
title: "meox1 dotplot panels (cdh5, cell-cycle/p53 panel)"
subtitle: "Kobayashi et al. 2026"
author: "Tyrone Chen"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    df_print: paged
    highlight: textmate
    keep_md: TRUE
    number_sections: TRUE
    theme: flatly
    toc: TRUE
    toc_float: TRUE
    toc_title: merged samples
editor_options:
  chunk_output_type: console
params:
  sample_index: 1
  sample_name: "default"
---

# README

Two small standalone dotplot panels, split out of the L3 Level dataset by `L3_celltype` x `Genotype`. Both are
one-off gene-panel additions (endothelial marker `cdh5`; a cell-cycle/p53 panel) rather than part of the main GO
pipeline, hence kept separate from `meox1_dotplot_goi.Rmd`.

## Outputs

Both under `../output/figure_extended/dotplots/`:

- `meox1_dataset_Level_03_L3_celltype_genotype_cdh5.pdf`
- `meox1_dataset_Level_03_L3_celltype_genotype_cellcyclelec_cdkn1a_tp53_cdkn1bb.pdf`

## Initial setup

```{r class.source="fold-hide", label="init", eval=TRUE}
library(ggplot2)
library(Seurat)
library(qs2)

outfile_dir <- "../output/figure_extended/dotplots/"
dir.create(outfile_dir, recursive = TRUE, showWarnings = FALSE)
infile_path <- "../../Saki_data/paper/D10051_meox1_dataset_Level_03_annotated.qs2"

rotate_x_90 <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

# standardised settings across dataset
genotype_colours <- c(
  "#dbe2c6", #wildtype
  "#657c95" #mutant
)
expression_colours <- c("#d9d9d9", "#40004b")
cols_cellcycle <- c( '#ffff66', '#cc85ff','#a0d0e0')
names(cols_cellcycle) <- c("G1/G0", "S", "G2M")

level_03_celltype_genotype_order <- c(
    "LEC_wildtype",
    "LEC_meox1_mutant",
    "pre_muLEC_wildtype",
    "pre_muLEC_meox1_mutant",
    "hmVEC_wildtype",
    "hmVEC_meox1_mutant",
    "mVEC_wildtype",
    "mVEC_meox1_mutant"
)

level_03 <- qs_read(infile_path)

level_03_relevel <- level_03
level_03_relevel$L3_celltype_genotype <- paste0(
    level_03_relevel$L3_celltype, "_", level_03_relevel$Genotype
    )
level_03_relevel <- SetIdent(
    level_03_relevel, value="L3_celltype_genotype"
    )
Idents(level_03_relevel) <- factor(
    level_03_relevel@active.ident, rev(
        level_03_celltype_genotype_order
        )
    )
```

## cdh5 panel

Source: `meox1_add_cdh5_to_dotplot.R`.

```{r class.source="fold-hide", label="cdh5_dotplot", eval=TRUE, fig.height=4, fig.width=6}
features_cdh5 <- c(
    "cdh5"
    )

dotplot_cdh5 <- DotPlot(
    object = level_03_relevel,
    features = features_cdh5,
    cluster.idents = F,
    cols = expression_colours
    ) + rotate_x_90 + ylab("") + xlab("")

print(dotplot_cdh5)

ggsave(
    paste0(outfile_dir, "meox1_dataset_Level_03_L3_celltype_genotype_cdh5.pdf"),
    dotplot_cdh5,
    device="pdf",
    width=6,
    height=4
)
```

## Cell-cycle / p53 panel

Source: `meox1_add_genes_to_dotplot.R`. Adds `cdkn1a`, `tp53`, `cdkn1bb` to the existing LEC-transition /
cell-cycle marker panel (`prox1a`, `tbx1`, `cdh6`, `pcna`, `mki67`).

```{r class.source="fold-hide", label="cellcycle_dotplot", eval=TRUE, fig.height=4, fig.width=6}
features_cellcycle <- c(
    "prox1a", "tbx1", "cdh6", "pcna", "mki67", "cdkn1a", "tp53", "cdkn1bb"
    )

dotplot_cellcycle <- DotPlot(
    object = level_03_relevel,
    features = features_cellcycle,
    cluster.idents = F,
    cols = expression_colours
    ) + rotate_x_90 + ylab("") + xlab("")

print(dotplot_cellcycle)

ggsave(
    paste0(outfile_dir, "meox1_dataset_Level_03_L3_celltype_genotype_cellcyclelec_cdkn1a_tp53_cdkn1bb.pdf"),
    dotplot_cellcycle,
    device="pdf",
    width=6,
    height=4
)
```

## For developers

Sample run command:

```{bash, eval=FALSE, class.source="fold-hide"}
Rscript -e "
rmarkdown::render(
  'meox1_dotplot_panels.Rmd',
  output_file = './meox1_dotplot_panels.html'
)
"
```
