---
title: "Explore new DEG approaches for yap1 dataset (Level 03)"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

# Introduction

Explore two key DEG options (mutant vs wildtype ):

- all VEC 

- combine iVECs with mVEC + combine iVEC with hmVEC

## Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)

})
```


## Load data

```{r}
level_03 <- qs_read(here('output/data/Datsets/paper/D10025_yap1_dataset_Level_03_annotated.qs2'))
```


## Load gene sets + paper settings 

```{r}
source(here("code/scRNAseq/general_scripts/yap1_meox1_important_genes.R"))
source(here("code/scRNAseq/general_scripts/yap1_meox1_paper_settings.R"))

names(genotype_colours) <- c("wildtype", "yap1_mutant")
```

## Plotting orders

```{r}

level_03_celltpe_order <- c(
  "LEC",
  "preLEC",
  "hmVEC",
  "mVEC",
  "cVEC",
  "iVEC"
)

```



# mutant vs WT in all VECs


```{r}
Idents(level_03) <- "L3_celltype"
seurat_subset <- subset(level_03, idents = c("mVEC", "hmVEC", "cVEC", "iVEC"))
seurat_subset$Genotype <- factor(seurat_subset$Genotype, levels = c("wildtype", "yap1_mutant"))
Idents(seurat_subset) <- "Genotype"

all_vec <- FindMarkers(object = seurat_subset,
                  group.by = "Genotype",
                  ident.1 = "yap1_mutant",
                  ident.2 = "wildtype",
                  min.pct = 0.1,
                  assay = "RNA",
                  logfc.threshold = 0) %>%
        rownames_to_column("Gene")
  

all_vec %>% filter( Gene == "meox1")

```

meox1 is not significant. 


# mutant vs WT in all LECs


```{r}
Idents(level_03) <- "L3_celltype"
seurat_subset <- subset(level_03, idents = c("LEC", "preLEC"))
seurat_subset$Genotype <- factor(seurat_subset$Genotype, levels = c("wildtype", "yap1_mutant"))
Idents(seurat_subset) <- "Genotype"

all_lec <- FindMarkers(object = seurat_subset,
                  group.by = "Genotype",
                  ident.1 = "yap1_mutant",
                  ident.2 = "wildtype",
                  min.pct = 0.1,
                  assay = "RNA",
                  logfc.threshold = 0) %>%
        rownames_to_column("Gene")
  

all_lec %>% filter( Gene == "meox1")

```



# mutant vs WT in iVEC+hmVEC


```{r}
Idents(level_03) <- "L3_celltype"
seurat_subset <- subset(level_03, idents = c("hmVEC", "iVEC"))
seurat_subset$Genotype <- factor(seurat_subset$Genotype, levels = c("wildtype", "yap1_mutant"))
Idents(seurat_subset) <- "Genotype"

ivec_hmVEC <- FindMarkers(object = seurat_subset,
                  group.by = "Genotype",
                  ident.1 = "yap1_mutant",
                  ident.2 = "wildtype",
                  min.pct = 0.1,
                  assay = "RNA",
                  logfc.threshold = 0) %>%
        rownames_to_column("Gene")
  

ivec_hmVEC %>% filter( Gene == "meox1")

```

meox1 is not significant. Not even raw p-value.


# mutant vs WT in iVEC+mVEC


```{r}
Idents(level_03) <- "L3_celltype"
seurat_subset <- subset(level_03, idents = c("mVEC", "iVEC"))
seurat_subset$Genotype <- factor(seurat_subset$Genotype, levels = c("wildtype", "yap1_mutant"))
Idents(seurat_subset) <- "Genotype"

ivec_mVEC <- FindMarkers(object = seurat_subset,
                  group.by = "Genotype",
                  ident.1 = "yap1_mutant",
                  ident.2 = "wildtype",
                  min.pct = 0.1,
                  assay = "RNA",
                  logfc.threshold = 0) %>%
        rownames_to_column("Gene")
  

ivec_mVEC %>% filter( Gene == "meox1")

```

meox1 is not significant. Not even raw p-value.


# reference - old DEG results

```{r}
all_degs <- read_csv(here("output/analysis/D10025_yap1_dataset/DEGs/Level_03_DEG_mutVSwt_L3_celltype_fc0.00_minpct_0.01.csv"), show_col_types = F)

#add new column names to avoid confusion
all_degs <- all_degs %>% 
  mutate(., `exp_wildtype[%]` = pct.2*100,
            `exp_mutant[%]` = pct.1*100)
  
  
```


```{r}
DT::datatable(all_degs %>% filter(., Gene=="meox1") %>% select(-c("pct.1", "pct.2")), rownames = FALSE)
```


Make a dotplot (to get info) for meox1 split by both celltype and genotype.


```{r, fig.align='center'}
level_03$L3_celltype_genotype <- paste0(level_03$L3_celltype, "_", level_03$Genotype)

dotplot <- DotPlot(level_03, features = "meox1", scale = F, group.by = "L3_celltype_genotype")

dotplot
```





