---
title: "Make figures for yap1 dataset"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

# Introduction

Make figures for all levels of the yap1 dataset. 

## Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(chisq.posthoc.test)

})
```


## Load data

```{r}
level_01 <- qs_read(here('output/data/Datsets/paper/D10025_yap1_dataset_Level_01_annotated.qs2'))
level_02 <- qs_read(here('output/data/Datsets/paper/D10025_yap1_dataset_Level_02_annotated.qs2'))
level_03 <- qs_read(here('output/data/Datsets/paper/D10025_yap1_dataset_Level_03_annotated.qs2'))
```

load mixing metric as well

```{r}
mixing_metric <- qs_read(here('output/analysis/D10025_yap1_dataset/MixingMetric/D10025_yap1_dataset_Level_03_mixing_metric.qs2'))
```

load DEGs

```{r}
hippo_gene_degs <- read_csv(here("output/analysis/D10025_yap1_dataset/DEGs/Level_03_DEG_hippo_targets_mutVSwt_L3_celltype_fc0.00_minpct_0.00.csv"), show_col_types = F)

all_degs <- read_csv(here("output/analysis/D10025_yap1_dataset/DEGs/Level_03_DEG_mutVSwt_L3_celltype_fc0.00_minpct_0.01.csv"), show_col_types = F)

all_markers <- read_csv(here("output/analysis/D10025_yap1_dataset/Markers/Level_03_Markers_L3_celltype_fc0.00_minpct_0.01.csv"), show_col_types = F)

```

load GO

```{r}
go_terms <- qs_read(here("output/analysis/D10025_yap1_dataset/enrichR/D10025_yap1_dataset_markers_L3_celltype_top100_enrichR.qs2"))
```


## Load gene sets + paper settings 

```{r}
source(here("code/scRNAseq/general_scripts/yap1_meox1_important_genes.R"))
source(here("code/scRNAseq/general_scripts/yap1_meox1_paper_settings.R"))

names(genotype_colours) <- c("wildtype", "yap1_mutant")
```

## Plotting orders

```{r}
level_01_order <- c(
  "Endothelial",
  "Endocardium",
  "Hematopoietic",
  "Erythrocyte",
  "Immune",
  "Neuronal",
  "Fibroblast",
  "Epithelial_Fibroblast",
  "Epithelial",
  "Eye_lens",
  "Chromatophore")

level_02_id_order <- c(
  "LEC_01",
  "VEC_01",
  "VEC_02",
  "VEC_03",
  "AEC_01",
  "AEC_02",
  "AEC_03",
  "AEC_04",
  "AEC_05",
  "cycling_EC_01",
  "Endocardium_01"
)

level_02_celltpe_order <- c(
  "LEC",
  "VEC",
  "AEC",
  "cycling_EC",
  "Endocardium"
)

level_03_celltpe_order <- c(
  "LEC",
  "preLEC",
  "hmVEC",
  "mVEC",
  "cVEC",
  "iVEC"
)



```


## Plotting colours 

```{r}
level_02_id_colours <- c(
  "#a1d99b",
  "#6baed6",
  "#4292c6",
  "#2171b5",
  "#fee0d2",
  "#fcbba1",
  "#fc9272",
  "#fb6a4a",
  "#ef3b2c",
  "#ffffd4",
  "#d9d9d9")

names(level_02_id_colours) <- level_02_id_order

level_02_celltype_colours <- c(
  "#a1d99b",
  "#6baed6",
  "#fc9272",
  "#ffffd4",
  "#d9d9d9")

names(level_02_celltype_colours) <- level_02_celltpe_order

level_03_celltype_colours <- c(
  "#41ae76",
  "#a1d99b",
  "#9ecae1",
  "#6baed6",
  "#4292c6",
  "#084594")

names(level_03_celltype_colours) <- level_03_celltpe_order

```


## Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## directories

```{r}
save_dir <- here("output/figures")
name <- "yap1_dataset"
```



# LEVEL 01 - ALL CELLS{.tabset}

```{r}
name_current <- sprintf("%s_Level_01", name)
```


n cells 

```{r}
level_01 %>% ncol()
```

n cells per genotype

```{r}
level_01$Genotype %>% table()
```


## UMAPs

Level 01 UMAPs don't have a colour scheme.

```{r, fig.align='center', fig.height=5, fig.width=7}
#clusterID
make_dim_umap(seurat = level_01, 
              group = "L1_cluster_id",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current)
```

```{r, fig.align='center', fig.height=5, fig.width=7}
#celltype
make_dim_umap(seurat = level_01, 
              group = "L1_celltype",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current)
```



```{r, fig.align='center', fig.height=5, fig.width=7}
#genotype
make_dim_umap(seurat = level_01, 
              group = "Genotype",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current, 
              colours = genotype_colours)
```

## Dotplot

```{r, fig.align='center', fig.height=6, fig.width=6}
level_01_relevel <- level_01
level_01_relevel 		<- SetIdent(level_01_relevel, value="L1_celltype")
Idents(level_01_relevel) <- factor(level_01_relevel@active.ident, level_01_order)

dotplot <- DotPlot(object = level_01_relevel,
                   features = level_01_annotation_genes,
                   cluster.idents = F, 
                   cols = expression_colours) + RotatedAxis() + coord_flip() + rotate_x + ylab("") + xlab("")

filename <- sprintf('%s/%s_L1_celltype_annotation.pdf', save_dir, name_current)
ggsave(filename,
       dotplot,
       device="pdf",
       width=5,
       height=7 )

dotplot
```

## Barplot

```{r, fig.align='center'}
barplot <- level_01@meta.data %>% 
  mutate(., Genotype = factor(Genotype, levels = names(genotype_colours)),
            L1_celltype = factor(L1_celltype, levels =level_01_order) )%>% 
  group_by(Genotype, L1_celltype, .drop=FALSE) %>% 
  summarise(., count =n()) %>% 
  ggplot(., aes(x =L1_celltype, y = count, fill = Genotype)) +
  geom_bar(stat="identity", position = "fill") +
  plot_sizing + rotate_x +
  scale_fill_manual(values = genotype_colours)+
  ylab("proportion") + xlab("")

filename <- sprintf('%s/%s_L1_celltype_genotype_barplot.pdf', save_dir, name_current)
ggsave(filename,
       barplot,
       device="pdf",
       width=7,
       height=4)

barplot
```






# LEVEL 02 - ALL ENDOTHELIAL CELLS{.tabset}

```{r}
name_current <- sprintf("%s_Level_02", name)
```


n cells 

```{r}
level_02 %>% ncol()
```

n cells per genotype

```{r}
level_02$Genotype %>% table()
```


## UMAPs

```{r, fig.align='center', fig.height=5, fig.width=7}
level_02$L2_cluster_id <- factor(x = level_02$L2_cluster_id,levels = level_02_id_order)

#clusterID
make_dim_umap(seurat = level_02, 
              group = "L2_cluster_id",
              width = 5, height = 5,
              path = save_dir,
              colours = level_02_id_colours,
              name = name_current)
```

```{r, fig.align='center', fig.height=5, fig.width=7}
level_02$L2_celltype <- factor(x = level_02$L2_celltype,levels = level_02_celltpe_order)

#celltype
make_dim_umap(seurat = level_02, 
              group = "L2_celltype",
              width = 5, height = 5,
              path = save_dir, 
              colours = level_02_celltype_colours,
              name = name_current)

```


```{r, fig.align='center', fig.height=5, fig.width=7}
#genotype
make_dim_umap(seurat = level_02, 
              group = "Genotype",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current, 
              colours = genotype_colours)
```

## Dotplot

```{r, fig.align='center', fig.height=5, fig.width=5}
level_02_relevel <- level_02
level_02_relevel 		<- SetIdent(level_02_relevel, value="L2_cluster_id")
Idents(level_02_relevel) <- factor(level_02_relevel@active.ident, level_02_id_order)

dotplot <- DotPlot(object = level_02_relevel,
                   features = level_02_annotation_genes,
                   cluster.idents = F, 
                   cols = expression_colours) + RotatedAxis() + coord_flip() + rotate_x + ylab("") + xlab("")

filename <- sprintf('%s/%s_L2_cluster_id_annotation.pdf', save_dir, name_current)
ggsave(filename,
       dotplot,
       device="pdf",
       width=5,
       height=5)

dotplot
```

## Barplot

```{r, fig.align='center'}
barplot <- level_02@meta.data %>% 
  mutate(., Genotype = factor(Genotype, levels = names(genotype_colours)),
            L2_cluster_id = factor(L2_cluster_id, levels =level_02_id_order) )%>% 
  group_by(Genotype, L2_cluster_id, .drop=FALSE) %>% 
  summarise(., count =n()) %>% 
  ggplot(., aes(x =L2_cluster_id, y = count, fill = Genotype)) +
  geom_bar(stat="identity", position = "fill") +
  plot_sizing + rotate_x +
  scale_fill_manual(values = genotype_colours)+
  ylab("proportion") + xlab("")

filename <- sprintf('%s/%s_L2_cluster_id_genotype_barplot.pdf', save_dir, name_current)
ggsave(filename,
       barplot,
       device="pdf",
       width=7,
       height=4)

barplot
```







# LEVEL 03 - VEC AND LEC{.tabset}

```{r}
name_current <- sprintf("%s_Level_03", name)
```


n cells 

```{r}
level_03 %>% ncol()
```

n cells per genotype

```{r}
level_03$Genotype %>% table()
```


## UMAPs


```{r, fig.align='center', fig.height=5, fig.width=7}
level_03$L3_celltype <- factor(x = level_03$L3_celltype,levels = level_03_celltpe_order)

#celltype
make_dim_umap(seurat = level_03, 
              group = "L3_celltype",
              width = 5, height = 5,
              path = save_dir, 
              colours = level_03_celltype_colours,
              name = name_current)

```


```{r, fig.align='center', fig.height=5, fig.width=7}
#genotype
make_dim_umap(seurat = level_03, 
              group = "Genotype",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current, 
              colours = genotype_colours)
```

## UMAPs - gene expression

```{r, fig.align='center', fig.height=10, fig.width=15}
genes_to_plot <-c("pcna", "mki67", "cdh6", "ccn2a", "ccn1", "prox1a")

plot_list <- lapply(genes_to_plot, function(gene){
  make_feature_umap(seurat = level_03, 
                  feature = gene,
              width = 5, height = 5,
              path = save_dir, 
              name = name_current)
})

wrap_plots(plot_list, ncol = 3)

```

## UMAPs - gene scores 

```{r, fig.align='center', fig.height=10, fig.width=10}
genescores_to_plot <-c("hippo_1", "mapk_1", "wnt_1", "ap1_1")

plot_list <- lapply(genescores_to_plot, function(genescore){
  make_genescore_umap(seurat = level_03, 
                  feature = genescore,
              width = 5, height = 5,
              path = save_dir, 
              name = name_current)
})

wrap_plots(plot_list, ncol = 2)

```

Split by genotype

```{r, fig.align='center', fig.height=10, fig.width=15}
genescores_to_plot <-c("hippo_1", "mapk_1", "wnt_1", "ap1_1")

plot_list <- lapply(genescores_to_plot, function(genescore){
  make_genescore_split_umap(seurat = level_03, 
                  feature = genescore,
              width = 10, height = 5,
              path = save_dir,  
              split = "Genotype",
              name = name_current)
})

wrap_plots(plot_list, ncol = 2)

```

## UMAPs - transition scores (mature)

```{r, fig.align='center', fig.height=5, fig.width=10}
p1 <-   make_genescore_umap(seurat = level_03, 
                  feature = "EMBO_LEC_score_1",
              width = 5, height = 5,
              path = save_dir,  
              colours = c("#d9d9d9", "darkgreen"),
              name = name_current)

p2 <- make_genescore_umap(seurat = level_03, 
                  feature = "EMBO_VEC_score_1",
              width = 5, height = 5,
              path = save_dir,  
              colours = c("#d9d9d9", "#406880"),
              name = name_current)

p1 + p2 + plot_layout(ncol = 2)
```

transition score with midpoint:

```{r, fig.align='center', fig.height=5, fig.width=7}
md <- level_03[[]]
coords <- Embeddings(level_03[["umap"]])
md <- cbind(md, coords)

midpoint <- mean(range(md$EMBO_transient_score))

# Order by absolute difference from the midpoint
md <- md[order(abs(md$EMBO_transient_score - midpoint)), ]

# plot
trans_score_legend <- ggplot(md, aes(x = umap_1, y = umap_2, color = EMBO_transient_score)) +
  geom_point(size = 1) + 
  scale_color_gradient2(low = "#406880", high = "darkgreen", mid = "azure2") + #not keeping original green bc its too light
  # viridis::scale_fill_viridis(option = "E", direction = 1) +
  theme_classic() 
trans_score_legend

umapFileName <- sprintf('%s/%s_UMAP_EMBO_transitionscore.pdf', save_dir, name_current)
ggsave(umapFileName,
       trans_score_legend,
       device="pdf",
       width=7,
       height=7)

trans_score_nolegend <- trans_score_legend + NoLegend() + NoAxes() + theme(plot.title = element_blank())
umapFileName <- sprintf('%s/%s_UMAP_EMBO_transitionscore_STRIPPED.pdf', save_dir, name_current)
ggsave(umapFileName,
       trans_score_nolegend,
       device="pdf",
       width=5,
       height=5)
```


## UMAPs - transition scores (developmental)

```{r, fig.align='center', fig.height=5, fig.width=10}
p1 <-   make_genescore_umap(seurat = level_03, 
                  feature = "LEC_dev_score_1",
              width = 5, height = 5,
              path = save_dir,  
              colours = c("#d9d9d9", "darkgreen"),
              name = name_current)

p1 + p2 + plot_layout(ncol = 2)
```

transition score:

```{r, fig.align='center', fig.height=5, fig.width=7}
md <- level_03[[]]
coords <- Embeddings(level_03[["umap"]])
md <- cbind(md, coords)

midpoint <- mean(range(md$EMBO_dev_transient_score))

# Order by absolute difference from the midpoint
md <- md[order(abs(md$EMBO_dev_transient_score - midpoint)), ]

# plot
trans_score_legend <- ggplot(md, aes(x = umap_1, y = umap_2, color = EMBO_dev_transient_score)) +
  geom_point(size = 1) + 
  scale_color_gradient2(low = "#406880", high = "darkgreen", mid = "azure2") + #not keeping original green bc its too light
  # viridis::scale_fill_viridis(option = "E", direction = 1) +
  theme_classic() 
trans_score_legend

umapFileName <- sprintf('%s/%s_UMAP_EMBO_dev_transitionscore.pdf', save_dir, name_current)
ggsave(umapFileName,
       trans_score_legend,
       device="pdf",
       width=7,
       height=7)

trans_score_nolegend <- trans_score_legend + NoLegend() + NoAxes() + theme(plot.title = element_blank())
umapFileName <- sprintf('%s/%s_UMAP_EMBO_transitionscore_STRIPPED.pdf', save_dir, name_current)
ggsave(umapFileName,
       trans_score_nolegend,
       device="pdf",
       width=5,
       height=5)
```

Alternatively, just highlight the transitions score. 

```{r, fig.align='center', fig.height=5, fig.width=7}

make_genescore_umap(seurat = level_03, 
                  feature = "EMBO_dev_transient_score",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current)
```




## UMAPs - mixing metric

```{r, fig.align='center', fig.height=5, fig.width=7}
# add to object
mixing_metric <- mixing_metric[names(mixing_metric) != "iVEC"] #this cluster is only mutant, no mixing
new_meta <- lapply(mixing_metric, mean) %>% unlist() %>% as.data.frame() %>% 
  rownames_to_column("L3_celltype") %>% dplyr::rename(mixing_metric=".")  %>% 
  full_join(level_03@meta.data, .)
rownames(new_meta) <- rownames(level_03@meta.data)
level_03@meta.data <- new_meta

# on UMAP
make_feature_umap(seurat = level_03, 
                  feature = "mixing_metric",
              width = 5, height = 5,
              path = save_dir, 
              name = name_current)
  
```


## Dotplot

```{r, fig.align='center', fig.height=5, fig.width=5}
level_03_relevel <- level_03
level_03_relevel 		<- SetIdent(level_03_relevel, value="L3_celltype")
Idents(level_03_relevel) <- factor(level_03_relevel@active.ident, level_03_celltpe_order)

dotplot <- DotPlot(object = level_03_relevel,
                   features = level_03_annotation_genes,
                   cluster.idents = F, 
                   cols = expression_colours) + RotatedAxis() + coord_flip() + rotate_x + ylab("") + xlab("")

filename <- sprintf('%s/%s_L3_celltype_annotation.pdf', save_dir, name_current)
ggsave(filename,
       dotplot,
       device="pdf",
       width=5,
       height=4)

dotplot
```

Hippo genes


```{r, fig.align='center', fig.height=5, fig.width=5}
level_03_relevel <- level_03
level_03_relevel 		<- SetIdent(level_03_relevel, value="L3_celltype")
Idents(level_03_relevel) <- factor(level_03_relevel@active.ident, level_03_celltpe_order)

dotplot <- DotPlot(object = level_03_relevel,
                   features = hippo_targets,
                   cluster.idents = F, 
                   cols = expression_colours) + RotatedAxis() + coord_flip() + rotate_x + ylab("") + xlab("")

filename <- sprintf('%s/%s_L3_celltype_hippo.pdf', save_dir, name_current)
ggsave(filename,
       dotplot,
       device="pdf",
       width=5,
       height=4)

dotplot
```


Hippo gene split by genotype

Add new order.

```{r}
level_03_celltpe_genotype_order <- c(
  "LEC_wildtype",
  "LEC_yap1_mutant",
  "preLEC_wildtype",
  "preLEC_yap1_mutant",
  "hmVEC_wildtype",
  "hmVEC_yap1_mutant",
  "mVEC_wildtype",
  "mVEC_yap1_mutant",
  "cVEC_wildtype",
  "cVEC_yap1_mutant",
  "iVEC_wildtype",
  "iVEC_yap1_mutant"
)
```


```{r, fig.align='center', fig.height=5, fig.width=7}
level_03_relevel <- level_03
level_03_relevel$L3_celltype_genotype <- paste0(level_03_relevel$L3_celltype, "_", level_03_relevel$Genotype)
level_03_relevel 		<- SetIdent(level_03_relevel, value="L3_celltype_genotype")
Idents(level_03_relevel) <- factor(level_03_relevel@active.ident, rev(level_03_celltpe_genotype_order))

dotplot <- DotPlot(object = level_03_relevel,
                   features = hippo_targets,
                   cluster.idents = F, 
                   cols = expression_colours)  + rotate_x + ylab("") + xlab("")

filename <- sprintf('%s/%s_L3_celltype_genotype_hippo.pdf', save_dir, name_current)
ggsave(filename,
       dotplot,
       device="pdf",
       width=7,
       height=4)

dotplot
```

Selected other genes, split by genoptype

```{r, fig.align='center', fig.height=5, fig.width=7}
level_03_relevel <- level_03
level_03_relevel$L3_celltype_genotype <- paste0(level_03_relevel$L3_celltype, "_", level_03_relevel$Genotype)
level_03_relevel 		<- SetIdent(level_03_relevel, value="L3_celltype_genotype")
Idents(level_03_relevel) <- factor(level_03_relevel@active.ident, rev(level_03_celltpe_genotype_order))

dotplot <- DotPlot(object = level_03_relevel,
                   features = c("prox1a", "tbx1", "cdh6", "pcna", "mki67"),
                   cluster.idents = F, 
                   cols = expression_colours)  + rotate_x + ylab("") + xlab("")

filename <- sprintf('%s/%s_L3_celltype_genotype_cellcyclelec.pdf', save_dir, name_current)
ggsave(filename,
       dotplot,
       device="pdf",
       width=6,
       height=4)

dotplot
```

## Dotplot - ratios

Just meox1

```{r, fig.align='center'}
input_ratio <- hippo_gene_degs %>% 
  filter(., Gene == "meox1") %>% 
  mutate(., group = factor(group, levels = level_03_celltpe_order))

make_ratio_dotplots(input_ratio,
                    width = 5, height = 3, 
                    name = paste0(name_current, "_meox1"), path = save_dir)
```

All hippo genes:


```{r, fig.align='center'}
input_ratio <- hippo_gene_degs %>% 
  mutate(., group = factor(group, levels = level_03_celltpe_order))

make_ratio_dotplots(input_ratio,
                    width = 4, height = 4, 
                    name = paste0(name_current, "_hippogenes"), path = save_dir)
```


## Barplot (and stats)

### Genotype vs celltype

```{r, fig.align='center'}
barplot <- level_03@meta.data %>% 
  mutate(., Genotype = factor(Genotype, levels = names(genotype_colours)),
            L3_celltype = factor(L3_celltype, levels =level_03_celltpe_order) )%>% 
  group_by(Genotype, L3_celltype, .drop=FALSE) %>% 
  summarise(., count =n()) %>% 
  ggplot(., aes(x =L3_celltype, y = count, fill = Genotype)) +
  geom_bar(stat="identity", position = "fill") +
  plot_sizing + rotate_x +
  scale_fill_manual(values = genotype_colours)+
  ylab("proportion") + xlab("")

filename <- sprintf('%s/%s_L3_celltype_genotype_barplot.pdf', save_dir, name_current)
ggsave(filename,
       barplot,
       device="pdf",
       width=6,
       height=4)

barplot
```

Stats to support this:

Using the chisq.posthoc.test package to perform a post-hoc analysis for this.

```{r}
M <-table(level_03$Genotype, level_03$L3_celltype) + 0.001 #add very small number to deal with test sensitivity to zeros

global_test <- chisq.test(M)
global_test$p.value


res <- chisq.posthoc.test(M) %>% as.data.frame() %>% 
    pivot_longer(., cols = level_03$L3_celltype %>% unique(), names_to = "L3_celltype", values_to = "entry") %>% 
    pivot_wider(., names_from = "Value", values_from = "entry")
  
res <- res %>% 
  distinct(., L3_celltype, .keep_all = T) %>%  #residuals for 2 groups are mirrored 
  mutate(., Residuals = abs(Residuals), 
            is_sig = case_when(`p values` > 0.05 ~ "", 
                              `p values` <= 0.0001 ~ "****",
                               `p values` <= 0.001 ~ "***",
                              `p values` <= 0.01 ~ "**",
                                `p values` <= 0.05 ~ "*"
                                 )) %>% 
  select(., -Dimension)

res %>% knitr::kable()

```


Add to barplot:

```{r, fig.align='center'}
barplot <- level_03@meta.data %>% 
  group_by(Genotype, L3_celltype, .drop=FALSE) %>% 
  summarise(., count =n()) %>% 
  full_join(., res) %>% 
    mutate(., Genotype = factor(Genotype, levels = names(genotype_colours)),
            L3_celltype = factor(L3_celltype, levels =level_03_celltpe_order) )%>% 
  mutate(., is_sig = ifelse(Genotype == "yap1_mutant", is_sig, NA)) %>% 
  ggplot(., aes(x =L3_celltype, y = count, fill = Genotype)) +
  geom_bar(stat="identity", position = "fill") +
  geom_text(aes(label = is_sig, y = 1.05),size=6) +
  plot_sizing + rotate_x +
  scale_fill_manual(values = genotype_colours)+
  ylab("proportion") + xlab("") + 
  ggtitle(paste0("chisq p-value: ",global_test$p.value))

filename <- sprintf('%s/%s_L3_celltype_genotype_barplot_stats.pdf', save_dir, name_current)
ggsave(filename,
       barplot,
       device="pdf",
       width=6,
       height=4)

barplot
```


### Genotype vs cell cycle stage (for each celltype)


```{r, fig.align='center', fig.height=5, fig.width=12}
#cell cycle
barplot <- level_03@meta.data %>% 
  mutate(., tricycle_stage = factor(tricycle_stage, levels = names(cols_cellcycle)),
            L3_celltype = factor(L3_celltype, levels =level_03_celltpe_order),
            Genotype = factor(Genotype, levels = names(genotype_colours))) %>% 
  group_by(tricycle_stage, L3_celltype,Genotype, .drop=FALSE) %>% 
  summarise(., count =n()) %>% 
  group_by(., Genotype, L3_celltype) %>% 
  mutate(., perc_pos = count/sum(count),
         perc_label = round(perc_pos, 2)*100) %>% 
  ggplot(., aes(x =Genotype, y = count, fill = tricycle_stage)) +
  geom_bar(stat="identity", position = "fill") +
  geom_text(aes(label = perc_label, y = perc_pos), position = position_stack(vjust = 0.5), size=5)+ 
  plot_sizing + rotate_x +
  scale_fill_manual(values = cols_cellcycle)+
  ylab("proportion") + xlab("") +
  facet_wrap(~L3_celltype, nrow=1) + strip_fix

filename <- sprintf('%s/%s_L3_celltype_tricycle_genotype_barplot.pdf', save_dir, name_current)
ggsave(filename,
       barplot,
       device="pdf",
       width=12,
       height=5)

barplot
```


Stats to support this:

```{r}
global_stats <- lapply(c("mVEC","preLEC", "LEC","hmVEC" ,"cVEC"), function(x){
  filtered_seurat <- level_03@meta.data %>% 
  filter(., L3_celltype == x) %>% # for each cell type population
  select(., Genotype, tricycle_stage) 
  M <-table(filtered_seurat$Genotype, filtered_seurat$tricycle_stage) + 0.001 #add very small number to deal with test sensitivity to zeros
  Xsq <- chisq.test(M)  # Prints test summary
  df_return <- data.frame(statistic = Xsq$statistic, 
                          pvalue= Xsq$p.value)
  return(df_return)
})

names(global_stats) <- c("mVEC","preLEC", "LEC","hmVEC" ,"cVEC")

global_stats <- global_stats %>% bind_rows(., .id = "L3_celltype")  %>% 
  mutate(., is_sig = case_when(pvalue > 0.05 ~ "", 
                              pvalue <= 0.0001 ~ "****",
                               pvalue <= 0.001 ~ "***",
                              pvalue <= 0.01 ~ "**",
                                pvalue<= 0.05 ~ "*"
                                 )) 

global_stats %>% knitr::kable()
```

Using the chisq.posthoc.test package to perform a post-hoc analysis for this. Residuals only end up being meaningful for the celltypes with global changes. 

```{r}
post_hoc_tests_to_run <-global_stats %>% filter(., pvalue < 0.05) %>% pull(L3_celltype)

all_x_tests <- lapply(post_hoc_tests_to_run, function(x){
  filtered_seurat <- level_03@meta.data %>% 
  filter(., L3_celltype == x) %>% # for each cell type population
  select(., Genotype, tricycle_stage) 
  M <-table(filtered_seurat$Genotype, filtered_seurat$tricycle_stage) + 0.001 #add very small number to deal with test sensitivity to zeros
  res <- chisq.posthoc.test(M) %>% as.data.frame() %>% 
    pivot_longer(., cols = level_03$tricycle_stage %>% unique(), names_to = "tricycle_stage", values_to = "entry") %>% 
    pivot_wider(., names_from = "Value", values_from = "entry")

  return(res)
})

names(all_x_tests) <- post_hoc_tests_to_run

all_x_tests_df <- all_x_tests %>% bind_rows(., .id = "L3_celltype") %>% 
  distinct(., L3_celltype, tricycle_stage, .keep_all = T) %>%  #residuals for 2 groups are mirrored 
  mutate(., Residuals = abs(Residuals), 
            is_sig = case_when(`p values` > 0.05 ~ "", 
                              `p values` <= 0.0001 ~ "****",
                               `p values` <= 0.001 ~ "***",
                              `p values` <= 0.01 ~ "**",
                                `p values` <= 0.05 ~ "*"
                                 )) %>% 
  select(., -Dimension)

all_x_tests_df %>% knitr::kable()

```

Barplot with stats:

```{r, fig.align='center', fig.height=5, fig.width=12}
#cell cycle
barplot <- level_03@meta.data %>% 
  group_by(tricycle_stage, L3_celltype,Genotype, .drop=FALSE) %>% 
  summarise(., count =n()) %>% 
  group_by(., Genotype, L3_celltype) %>% 
  mutate(., perc_pos = count/sum(count),
         perc_label = round(perc_pos, 2)*100) %>% 
  ungroup() %>% 
  full_join(., all_x_tests_df) %>% 
  full_join(., global_stats %>% select(., is_sig, L3_celltype, pvalue) %>% dplyr::rename(is_sig_global=is_sig, pvalue_global=pvalue)) %>%  
  mutate(., tricycle_stage = factor(tricycle_stage, levels = names(cols_cellcycle)),
            L3_celltype = factor(L3_celltype, levels =level_03_celltpe_order),
            Genotype = factor(Genotype, levels = names(genotype_colours)), 
            is_sig = ifelse(Genotype == "yap1_mutant", is_sig, NA), 
            is_sig_global = ifelse(Genotype == "yap1_mutant", is_sig_global, NA),
            pvalue_global = ifelse(Genotype == "yap1_mutant" & is_sig_global !="", pvalue_global, NA)) %>% 
  ggplot(., aes(x =Genotype, y = count, fill = tricycle_stage)) +
  geom_bar(stat="identity", position = "fill") +
  geom_text(aes(label = perc_label, y = perc_pos), position = position_stack(vjust = 0.5), size=5)+ 
  geom_text(aes(label = is_sig, y = perc_pos), position = position_stack(vjust = 0.9), size=5)+ 
  geom_text(aes(label = is_sig_global, y = 1.1, x= 1.5), size=7)+ 
  geom_text(aes(label = pvalue_global, y = 1.05, x= 1.5), size=3)+ 
  plot_sizing + rotate_x +
  scale_fill_manual(values = cols_cellcycle)+
  ylab("proportion") + xlab("") +
  facet_wrap(~L3_celltype, nrow=1) + strip_fix

filename <- sprintf('%s/%s_L3_celltype_tricycle_genotype_barplot_stats.pdf', save_dir, name_current)
ggsave(filename,
       barplot,
       device="pdf",
       width=12,
       height=5)

barplot
```



## Volcano plot

Volcano plot with log2FC.

```{r, fig.align='center', fig.height=5, fig.width=7}
degs_mark <- all_degs %>% 
  filter(., group == "hmVEC" & p_val_adj < 0.05) 

hippo_mark <- all_degs %>% 
  filter(., group == "hmVEC" & Gene %in% hippo_targets)

plot <- all_degs %>% 
  filter(., group == "hmVEC") %>% 
  ggplot(., aes(x = avg_log2FC, y = -log10(p_val))) +
  geom_point(colour="#d9d9d9") +
  geom_point(data = degs_mark, colour="#969696") +
  geom_point(data = hippo_mark, colour="red", size=2) +
  ggrepel::geom_text_repel(data = hippo_mark, aes(label= Gene), size =5) +
  plot_sizing + ylab("-log10(p-value)") + xlab("log2FC")


filename <- sprintf('%s/%s_hmVEC_volcano_adjpvalue.pdf', save_dir, name_current)
ggsave(filename,
       plot,
       device="pdf",
       width=6,
       height=5)

plot
```

Just p-value

```{r, fig.align='center', fig.height=5, fig.width=7}
degs_mark <- all_degs %>% 
  filter(., group == "hmVEC" & p_val < 0.05) 

hippo_mark <- all_degs %>% 
  filter(., group == "hmVEC" & Gene %in% hippo_targets)

plot <- all_degs %>% 
  filter(., group == "hmVEC") %>% 
  ggplot(., aes(x = avg_log2FC, y = -log10(p_val))) +
  geom_point(colour="#d9d9d9") +
  geom_point(data = degs_mark, colour="#969696") +
  geom_point(data = hippo_mark, colour="red", size=2) +
  ggrepel::geom_text_repel(data = hippo_mark, aes(label= Gene), size =5) +
  plot_sizing + ylab("-log10(p-value)") + xlab("log2FC")


filename <- sprintf('%s/%s_hmVEC_volcano_pvalue.pdf', save_dir, name_current)
ggsave(filename,
       plot,
       device="pdf",
       width=6,
       height=5)

plot
```


## DEG barplots

p-value

```{r, fig.align='center', fig.height=4, fig.width=15}
plot <- hippo_gene_degs %>% 
  mutate(., group = factor(group, levels = level_03_celltpe_order)) %>% 
  mutate(., label = case_when(p_val > 0.05 ~ "", 
                              p_val <= 0.0001 ~ "****",
                               p_val <= 0.001 ~ "***",
                              p_val <= 0.01 ~ "**",
                                 p_val <= 0.05 ~ "*"
                                 )) %>% 
  ggplot(., aes(x = Gene, y = avg_log2FC, fill = group)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label=label)) +
  plot_sizing + rotate_x +
  scale_fill_manual(values = level_03_celltype_colours) +
  facet_wrap(~group, nrow=1) + strip_fix + ylab("log2FC") + xlab("")


filename <- sprintf('%s/%s_DEG_hippo_barplot_pvalue.pdf', save_dir, name_current)
ggsave(filename,
       plot,
       device="pdf",
       width=12,
       height=4)

plot
```

adjusted

```{r, fig.align='center', fig.height=4, fig.width=15}
plot<- hippo_gene_degs %>% 
  mutate(., group = factor(group, levels = level_03_celltpe_order)) %>% 
  mutate(., label = case_when(p_val_adj > 0.05 ~ "", 
                              p_val_adj <= 0.0001 ~ "****",
                               p_val_adj <= 0.001 ~ "***",
                              p_val_adj <= 0.01 ~ "**",
                                 p_val_adj <= 0.05 ~ "*"
                                 )) %>% 
  ggplot(., aes(x = Gene, y = avg_log2FC, fill = group)) +
  geom_bar(position = position_dodge2(), stat = "identity") +
  geom_text(aes(label=label)) +
  plot_sizing + rotate_x +
  scale_fill_manual(values = level_03_celltype_colours) +
  facet_wrap(~group, nrow=1) + strip_fix+ ylab("log2FC") + xlab("")

filename <- sprintf('%s/%s_DEG_hippo_barplot_adjustedpvalue.pdf', save_dir, name_current)
ggsave(filename,
       plot,
       device="pdf",
       width=12,
       height=4)

plot
```


## GO plot for IL_VEC

Export only iVEC results for GO analysis. 

```{r}
export_sub <- go_terms %>% 
  mutate(., Term = gsub("_WP.*", "", Term)) %>% 
  filter(., cluster == "iVEC")  %>% 
  mutate(Overlap = gsub("/", " out of ", Overlap)) #because excel keeps thinking this is a date

filename <- sprintf("%s/%s_iVEC_GO_WP_results.csv", save_dir, name_current)
write.csv(x = export_sub, file = filename, row.names = FALSE, quote = T)

```

Also export the genes that went into this analysis

```{r}
export_sub <-  all_markers %>% filter(., p_val_adj < 0.05 & cluster == "iVEC") %>%
    top_n(., n=100, wt = -avg_log2FC) 

filename <- sprintf("%s/%s_iVEC_markers_for_enrichment.csv", save_dir, name_current)
write.csv(x = export_sub, file = filename, row.names = FALSE, quote = T)

```

Make plot. 

```{r, fig.align='center', fig.height=5, fig.width=7}
plot <- go_terms %>% 
  mutate(., Term = gsub("_WP.*", "", Term)) %>% 
  filter(., cluster == "iVEC" & Adjusted.P.value < 0.05) %>% 
  ggplot(., aes(y = Genes, x = Combined.Score)) +
  geom_bar(stat="identity", fill = "purple4") +
  geom_text(aes(label = Term, x=0), colour = "grey90", size = 6, hjust=-0.01) +
  plot_sizing + xlab("enrichR score") 
  
filename <- sprintf('%s/%s_ivec_barplot_GOterms.pdf', save_dir, name_current)
ggsave(filename,
       plot,
       device="pdf",
       width=10,
       height=4)

plot

```

## Boxplot for scores

Normality test: KS only, shapiro can't handle > 5000 sample size. If the p-value is > 0.05 then distribution is significantly different from normal. 

Which means:

for pvalue < 0.05 -> non-parametric test (KW+wilcox)

for pvalue > 0.05 -> parametric test (ANOVA+ttest)

Check for normality for all scores:

```{r}
all_scores <- c("hippo_1", "mapk_1", "wnt_1", "ap1_1", "EMBO_LEC_score_1", "LEC_dev_score_1", 
                "EMBO_VEC_score_1", "EMBO_transient_score", "EMBO_dev_transient_score")

all_res <- lapply(all_scores, function(score){
  data_in <- level_03@meta.data %>% pull(score)
  df_out <- data.frame(ks_res = ks.test(data_in, "pnorm")$p.value)
  plots_out <- ggdensity(data_in) + ggqqplot(data_in) + plot_layout(ncol =1)+ plot_annotation(title = score)
  
  return(list(plot= plots_out, stats=df_out))
  
})
names(all_res) <- all_scores
```

Show normality plots

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(all_res, function(item) wrap_elements(item$plot))
wrap_plots(plot_list)
```

Show stats results

```{r}
lapply(all_res, function(item) item$stats) %>% bind_rows(., .id="score")
```

All are significant, so everything will be non-parametric! I think this is also related to the large number of cells..


Make boxplots with stats for everything:

```{r}
run_boxplot_stats_adjusted_pval <- function(df_in= level_03@meta.data,
                              score, 
                              celltype_order){
  
  #new labels for x axis
  new_labels <- df_in %>% 
  select(., score, L3_celltype, Genotype) %>% 
  mutate(., L3_celltype = factor(L3_celltype, levels = celltype_order)) %>% 
  group_by(Genotype, L3_celltype) %>% 
  summarise(., count = n()) %>% 
  group_by(L3_celltype) %>% 
  summarise(., label = paste0(L3_celltype,"\nwt=", count[1] , ", mut=",count[2])) %>% 
  pull(label) %>% unique()
  

  #rename
  colnames(df_in)[colnames(df_in) == score] <- "score"
  
  #y limits for test
  y_max <- df_in$score %>% max()
  y_pairwise <- y_max + (y_max*0.2)
  y_global <- y_max + (y_max*0.4)
  
  #get stats
  stat_list <- list()
  stat_list[["global"]] <- compare_means(score ~ Genotype,
                               data = df_in, 
                               method = "kruskal.test")
  stat_list[["pairwise"]] <- compare_means(score ~ Genotype,
                                 data =df_in, 
                                 group.by = "L3_celltype", 
                                 method = "wilcox.test") %>% 
    mutate(., sig_sign = case_when(p.adj > 0.05 ~ "ns", #add for adjusted
                                 p.adj <= 0.05 & p.adj > 0.01 ~ "*",
                                 p.adj <= 0.01 & p.adj > 0.001 ~ "**",
                                 p.adj <= 0.001 & p.adj > 0.0001 ~ "***",
                                 p.adj <= 0.0001 ~ "****"))
  
  plot <- df_in %>% 
    select(., score, L3_celltype, Genotype) %>% 
  mutate(., L3_celltype = factor(L3_celltype, levels = celltype_order),
            Genotype = factor(Genotype, levels = c("wildtype", "yap1_mutant"))) %>% 
  ggplot(.) +
  geom_boxplot(aes(x = L3_celltype, y = score, fill = Genotype), 
               position = position_dodge2(),outlier.shape = NA) +
  scale_fill_manual(values = genotype_colours) +
  theme_bw(base_size = 14)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15)) +
  xlab("") + ylab(paste0(score, " score")) +
  geom_jitter(aes(x = L3_celltype, y = score, fill = Genotype),
              size = 0.1, alpha = 0.5, 
              position  = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.75)) +
  geom_text(data = stat_list$pairwise, aes(x = L3_celltype, y = y_pairwise, label = sig_sign)) +
  geom_text(data = stat_list$global, aes(x = 2.5, y = y_global, label = paste0(method, ", p = ", p.adj))) +
  scale_x_discrete(labels= new_labels) +
  ggtitle(score)
  
  
  df_stats <- bind_rows(stat_list)
  
  return(list(plot=plot, stat=df_stats))

}


```


Run for all scores:

```{r, warning=FALSE, message=FALSE}
results <- lapply(all_scores, function(score_input) run_boxplot_stats_adjusted_pval(score = score_input, 
                                                                        celltype_order = level_03_celltpe_order))
```

```{r, fig.align='center', fig.height=15, fig.width=15}
plot_list <- lapply(results, function(item) item$plot)

patch <- wrap_plots(plot_list, guides = "collect")

filename <- sprintf('%s/%s_module_scores_boxplots.pdf', save_dir, name_current)
ggsave(filename,
       patch,
       device="pdf",
       width=15,
       height=15)

patch

```


Show stats details:

```{r}
names(results) <- all_scores
stats_DF <- lapply(results, function(item) item$stat) %>% bind_rows(., .id="score")

#also export 
filename <- sprintf("%s/%s_module_scores_boxplots_stats.csv", save_dir, name_current)
write.csv(x = stats_DF, file = filename, row.names = FALSE)

DT::datatable(stats_DF, class = 'cell-border stripe', rownames = FALSE)

```


Make plots with raw pvalue:

```{r}
run_boxplot_stats_RAW_pval <- function(df_in= level_03@meta.data,
                              score, 
                              celltype_order){
  
  #new labels for x axis
  new_labels <- df_in %>% 
  select(., score, L3_celltype, Genotype) %>% 
  mutate(., L3_celltype = factor(L3_celltype, levels = celltype_order)) %>% 
  group_by(Genotype, L3_celltype) %>% 
  summarise(., count = n()) %>% 
  group_by(L3_celltype) %>% 
  summarise(., label = paste0(L3_celltype,"\nwt=", count[1] , ", mut=",count[2])) %>% 
  pull(label) %>% unique()
  

  #rename
  colnames(df_in)[colnames(df_in) == score] <- "score"
  
  #y limits for test
  y_max <- df_in$score %>% max()
  y_pairwise <- y_max + (y_max*0.2)
  y_global <- y_max + (y_max*0.4)
  
  #get stats
  stat_list <- list()
  stat_list[["global"]] <- compare_means(score ~ Genotype,
                               data = df_in, 
                               method = "kruskal.test")
  stat_list[["pairwise"]] <- compare_means(score ~ Genotype,
                                 data =df_in, 
                                 group.by = "L3_celltype", 
                                 method = "wilcox.test")
  
  plot <- df_in %>% 
    select(., score, L3_celltype, Genotype) %>% 
  mutate(., L3_celltype = factor(L3_celltype, levels = celltype_order),
            Genotype = factor(Genotype, levels = c("wildtype", "yap1_mutant"))) %>% 
  ggplot(.) +
  geom_boxplot(aes(x = L3_celltype, y = score, fill = Genotype), 
               position = position_dodge2(),outlier.shape = NA) +
  scale_fill_manual(values = genotype_colours) +
  theme_bw(base_size = 14)+ 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=15)) +
  xlab("") + ylab(paste0(score, " score")) +
  geom_jitter(aes(x = L3_celltype, y = score, fill = Genotype),
              size = 0.1, alpha = 0.5, 
              position  = position_jitterdodge(jitter.width = 0.3, dodge.width = 0.75)) +
  geom_text(data = stat_list$pairwise, aes(x = L3_celltype, y = y_pairwise, label = p.signif)) +
  geom_text(data = stat_list$global, aes(x = 2.5, y = y_global, label = paste0(method, ", p = ", p.adj))) +
  scale_x_discrete(labels= new_labels) +
  ggtitle(paste0(score, ", RAW PVALUE"))
  
  
  return(plot)

}


```


Run for all scores:

```{r, warning=FALSE, message=FALSE}
plot_list <- lapply(all_scores, function(score_input) run_boxplot_stats_RAW_pval(score = score_input, 
                                                                        celltype_order = level_03_celltpe_order))
```

```{r, fig.align='center', fig.height=15, fig.width=15}
patch <- wrap_plots(plot_list, guides = "collect")

filename <- sprintf('%s/%s_module_scores_boxplots_RAWpvalue.pdf', save_dir, name_current)
ggsave(filename,
       patch,
       device="pdf",
       width=15,
       height=15)

patch

```







