---
title: "D10025 yap1 dataset: Annotate Level 01 (all cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Annotate Level 01 of the yap1 dataset


### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(tricycle)
  library(scGate)
})
```

### Load objects

Objects made here: ./pre_processing_D10025_yap1_dataset_makeObject_Level01_SCT.Rmd

Only loading merged ones for now.

```{r}
input_paths <- c(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_SCT_kept_doublets.qs2'),
                 here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_SCT.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

Give them names in the object

```{r}
list_objects$D10025_yap1_dataset_Level_01_SCT_kept_doublets@misc[["name"]] <- "Level_01_kept_doublets"
list_objects$D10025_yap1_dataset_Level_01_SCT@misc[["name"]] <- "Level_01"
```


Read in mouse -> fish names

```{r}
mouse_to_fish <- qs_read(here("data/genesets/geneMap_mmusculus_to_drerio.qs2"))
```


### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```

## Tricycle: Run cell cycle analysis

Make function.

```{r}

run_tricycle_for_fish <- function(seurat, x_species_dictionary = mouse_to_fish){
  #prepare
  sce <- as.SingleCellExperiment(seurat, assay = "RNA") #convert to SCE
  sub_sce <- sce[rownames(sce) %in% x_species_dictionary$drerio_zfin_id_symbol,] #subset to genes in dic
  sub_x_species_dictionary <- x_species_dictionary %>%  #subset dictionary to only include the genes that we kept 
  filter(., drerio_zfin_id_symbol %in% rownames(sub_sce)) %>% 
  distinct(drerio_zfin_id_symbol, .keep_all = TRUE) #needed for function
  sub_x_species_dictionary <- sub_x_species_dictionary[match(rownames(sub_sce),
                                                             sub_x_species_dictionary$drerio_zfin_id_symbol),]
  stopifnot(all(sub_x_species_dictionary$drerio_zfin_id_symbol == rownames(sub_sce))) #check
  
  #run tricycle
  sub_sce <- project_cycle_space(sub_sce, 
                                 gname = sub_x_species_dictionary$mmusculus_mgi_symbol, 
                                 species = "mouse", 
                                 gname.type = "SYMBOL")
  #estimate position
  sub_sce <- estimate_cycle_position(sub_sce)
  
  
  #add back to seurat object
  seurat$tricycle_position <- sub_sce$tricyclePosition
  
  #add categorical cell cycle:
  # https://github.com/lb15/multiciliation_cycle/blob/main/mcc_timecourse/mcc_timecourse_analysis.R#L619
  seurat$tricycle_stage <- case_when(seurat$tricycle_position > pi/2 & seurat$tricycle_position < pi ~ "S",
                             seurat$tricycle_position >= pi & seurat$tricycle_position < pi*7/4 ~ "G2M",
                             TRUE ~ "G1/G0")
  
  rm(sce,sub_sce, sub_x_species_dictionary); gc()
  return(seurat)
  

}

```


```{r}
list_objects <- lapply(list_objects, run_tricycle_for_fish)
```

### Visualise test genes

Write function:

```{r}

make_tricycle_plots <- function(seurat){
  #colours 
  cols_cellcycle <- c('#ffff66', '#a0d0e0', '#cc85ff')
  names(cols_cellcycle) <- c('G1', 'G2M', 'S')
  rects <- data.frame(xmin = c(0, 0.5*pi, pi, 1.75*pi),
                    xmax = c(0.5*pi, pi, 1.75*pi, 2*pi),
                    fill = c('#ffff66',  '#cc85ff','#a0d0e0', '#ffff66'))
  
  
  plot_list <- list()
  plot_list[[1]] <- fit_periodic_loess(seurat$tricycle_position,
                                       seurat[["RNA"]]@layers$data[which(rownames(seurat) == 'cdt1'),],
                                       plot = TRUE,
                                       x_lab = "Cell cycle position \u03b8", y_lab = "cdt1 expression",
                                       fig.title = paste0("Expression of cdt1 (G1-phase)"))$fig +
    plot_sizing +
    annotate(geom = "text", x = c(0.25*pi, 0.75*pi, 1.4*pi, 1.9*pi), y = 4, label = c("G0/G1", "S", "G2M","G0/G1"), size = 4) +
    coord_cartesian(expand = FALSE, clip = "off", ylim = c(0, 4.5)) +
    geom_vline(xintercept=0.5*pi,linetype="dashed")+
    geom_vline(xintercept =pi,linetype="dashed")+
    geom_vline(xintercept=1.75*pi,linetype="dashed")  +
    geom_rect(aes(xmin = xmin, xmax = xmax, fill = fill, ymin = -Inf, ymax = Inf),
              data = rects, inherit.aes = FALSE, size = 1, alpha = .1) +
    scale_fill_manual(values = c( '#a0d0e0', '#cc85ff', '#ffff66','#ffff66')) +
    theme(legend.position = "none")


  plot_list[[2]] <- fit_periodic_loess(seurat$tricycle_position,
                                       seurat[["RNA"]]@layers$data[which(rownames(seurat) == 'pcna'),],
                                       plot = TRUE,
                                       x_lab = "Cell cycle position \u03b8", y_lab = "pcna expression",
                                       fig.title = paste0("Expression of pcna (S-phase)"))$fig + 
    plot_sizing +
    annotate(geom = "text", x = c(0.25*pi, 0.75*pi, 1.4*pi, 1.9*pi), y = 4, label = c("G0/G1", "S", "G2M","G0/G1"), size = 4) +
    coord_cartesian(expand = FALSE, clip = "off", ylim = c(0, 6)) +
    geom_vline(xintercept=0.5*pi,linetype="dashed")+
    geom_vline(xintercept =pi,linetype="dashed")+
    geom_vline(xintercept=1.75*pi,linetype="dashed") +
    geom_rect(aes(xmin = xmin, xmax = xmax, fill = fill, ymin = -Inf, ymax = Inf),
              data = rects, inherit.aes = FALSE, size = 1, alpha = .1) +
    scale_fill_manual(values = c( '#a0d0e0', '#cc85ff', '#ffff66','#ffff66')) +
    theme(legend.position = "none")
  
  plot_list[[3]] <- fit_periodic_loess(seurat$tricycle_position,
                                       seurat[["RNA"]]@layers$data[which(rownames(seurat) == 'mki67'),],
                                       plot = TRUE,
                                       x_lab = "Cell cycle position \u03b8", y_lab = "mki67 expression",
                                       fig.title = paste0("Expression of mki67 (G2M-phase)"))$fig +
    plot_sizing +
    annotate(geom = "text", x = c(0.25*pi, 0.75*pi, 1.4*pi, 1.9*pi), y = 4, label = c("G0/G1", "S", "G2M","G0/G1"), size = 4) +
    coord_cartesian(expand = FALSE, clip = "off", ylim = c(0, 4.5)) +
    geom_vline(xintercept=0.5*pi,linetype="dashed")+
    geom_vline(xintercept =pi,linetype="dashed")+
    geom_vline(xintercept=1.75*pi,linetype="dashed") +
    geom_rect(aes(xmin = xmin, xmax = xmax, fill = fill, ymin = -Inf, ymax = Inf),
              data = rects, inherit.aes = FALSE, size = 1, alpha = .1) +
    scale_fill_manual(values = c( '#a0d0e0', '#cc85ff', '#ffff66','#ffff66')) +
    theme(legend.position = "none")
  
  return(plot_list)
}


```

Run for all

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, make_tricycle_plots)

wrap_plots(unlist(plot_list, recursive = F)) + plot_annotation(title = "top = with doublets, bottom = without")
```

### UMAPs for cell cycle stages

```{r}
run_umap_stages <- function(seurat){
  #colours 
  cols_cellcycle <- c('#ffff66', '#a0d0e0', '#cc85ff')
  names(cols_cellcycle) <- c('G1/G0', 'G2M', 'S')
  #plot
  DimPlot(seurat, group.by = "tricycle_stage", cols = cols_cellcycle) + ggtitle(seurat@misc$name)
}
```

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, run_umap_stages)

wrap_plots(plot_list, ncol = 2) + plot_layout(guides = "collect", axis_titles = "collect")
```



## Clustree

```{r, fig.align='center', fig.height=12, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  clustree(seurat, prefix = c("SCT_snn_res.")) + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2)
```


Resolution 0.4 looks pretty solid for both of them. This doesn't have to be very detailed yet, it is most important for us to isolate the endothlial cells. 

## Low quality clusters

Do any clusters have high mt and/or low count? Being mindful that some celltypes do have naturally higher %mt and counts.

```{r, fig.align='center', fig.height=8, fig.width=15}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nCount_SCT), x = SCT_snn_res.0.4, fill=SCT_snn_res.0.4)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[2]] <- seurat@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = SCT_snn_res.0.4, fill=SCT_snn_res.0.4)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  return(plot_list) 
})

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

Nothing very obvious.


## scGate: Find endothelial subset

Using scGate as a computational gating strategy to identify all endothelial cells. Endothelial cells are marked by expression of kdrl, and absence of expression of col1a1a (Fibroblasts), epcam (Epithelial), gata2b (haematopoietic), myod1 (Muscle). 

Mark clusters that have > 95% purity for resolution 0.5.

```{r}
endothelial_gating_model <- gating_model(name = "endothelial", signature = c("kdrl+", 
                                                                    "col1a1a-",
                                                                    "epcam-",
                                                                    "gata2b-", 
                                                                    "myod1-"))

#write function
run_scgate_endo <- function(seurat){
  seurat <- scGate(data = seurat,
                   model = endothelial_gating_model,
                   output.col.name = "is_endothlial")
  
  pure_clusters <- seurat@meta.data %>%
    group_by(is_endothlial, SCT_snn_res.0.4) %>%
    summarise(., count = n()) %>%
    group_by(SCT_snn_res.0.4) %>%
    summarise(., SCT_snn_res.0.4_cluster_purity = count[is_endothlial == "Pure"]/sum(count)) %>%
    full_join(seurat@meta.data, ., by = "SCT_snn_res.0.4")
  
  rownames(pure_clusters) <- rownames(seurat@meta.data)
  
  seurat@meta.data <- pure_clusters

  return(seurat)
}

```


```{r}
list_objects <- lapply(list_objects, run_scgate_endo)
```

Show UMAPs with a) is pure and b) cluster purity.

```{r}
run_umap_scgate_results <- function(seurat){
  #colours
  pure_colours <- c("grey", "green3"); names(pure_colours) <- c("Impure", "Pure")
  
  plot_list <- list()
  plot_list[[1]] <- DimPlot(seurat, group.by = "is_endothlial", cols = pure_colours) + ggtitle(sprintf("Endo Gate, %s", seurat@misc$name))
  plot_list[[2]] <- FeaturePlot(seurat, features = "SCT_snn_res.0.4_cluster_purity") + ggtitle(sprintf("Cluster Purity, %s", seurat@misc$name))
  
  return(plot_list)
}
```

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, run_umap_scgate_results)

wrap_plots(unlist(plot_list, recursive = F), ncol = 2) + plot_layout(guides = "collect", axis_titles = "collect")
```

Make barplots for endo purity

```{r, fig.align='center', fig.height=5, fig.width=12}
plot_list <- lapply(list_objects, function(seurat){
  seurat@meta.data %>% 
    group_by(., SCT_snn_res.0.4) %>% 
    summarise(SCT_snn_res.0.4_cluster_purity) %>% distinct() %>% 
    ggplot(., aes(x= reorder(SCT_snn_res.0.4, -SCT_snn_res.0.4_cluster_purity),  y = SCT_snn_res.0.4_cluster_purity))+
    geom_bar(stat = "identity") + plot_sizing +
    ylab("Endo Purity per cluster") + xlab("cluster") + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2) + plot_layout(guides = "collect", axis_titles = "collect", axes = "collect_y")

```

## Dotplot for genes

Use some known marker genes for celltypes to double check scGate annotation.

```{r}
dotplot.level.01 <- c(
  "cdh6",
  "prox1a",
  "stab2",
  "lyve1b",
  "mrc1a",
  "flt1",
  "dll4",
  "esm1",
  "hlx1",
  "hand2",
  "fn1a",
  "gata2b",
  "runx1",
  "alas2",
  "blvrb",
  "spi1b",
  "mpx",
  "neurod4",
  "nova2",
  "cdh1",
  "epcam",
  "myod1",
  "myog",
  "pmela",
  "dct",
  "lum",
  "col1a1a",
  "krt4",
  "cyt1",
  "cryba1b",
  "crybb1",
  "ednrba",
  "pnp4a",
  "gch2",
  "aox5")
```

Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(seurat@misc$name)  +plot_sizing_seurat + rotate_x
}

```

Make dotplots for all clusters

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, function(seurat) run_dotplot(seurat, dotplot.level.01, group = "SCT_snn_res.0.4"))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

Make dotplot for clusters that have mid-range purity.

```{r}
run_dotplot_mid_purity <- function(seurat, genes, limits){
  clusters <- seurat@meta.data %>% 
    filter(., SCT_snn_res.0.4_cluster_purity > limits[1] & SCT_snn_res.0.4_cluster_purity < limits[2]) %>% 
    pull(SCT_snn_res.0.4) %>% unique()
  
  seurat <- SetIdent(seurat, value = "SCT_snn_res.0.4")
  DotPlot(seurat, features = genes, group.by = "SCT_snn_res.0.4", idents = clusters) + ggtitle(seurat@misc$name)  +plot_sizing_seurat + rotate_x
}

```

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, function(seurat) run_dotplot_mid_purity(seurat, dotplot.level.01, limits = c(0.4, 0.8)))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

The clusters with mid-range endothelial cell compositions clearly express genes for other cell types, so we will remove them.


## Export 

Export them all. Selection of clusters to move to Level 02 will happen in next script

```{r}

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_SCT_annotated.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_01_SCT, file_name)

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_SCT_kept_doublets_annotated.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_01_SCT_kept_doublets , file_name)

```



