---
title: "Combined WT only dataset: Make Level 02 (all endothelial cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 02 object for combined wildtypes objects

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_combining_wildtypes_dataset_makeObject_Level01_annotation.Rmd

```{r}

level_01 <- qs_read(here('output/data/Datsets/merged_wildtypes/merged_wildtypes_Level_01_annotated.qs2'))

```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: Endothelial clusters only

Subset Level 01 to only include endothelial subsets. By going through ./pre_processing_combining_wildtypes_dataset_makeObject_Level01_annotation.Rmd, we can see that:

- there are no low quality clusters that stand out

- we can find endothelial subsets using a mix of marker genes and scGate endothelial purity (95%) for clusters


```{r}
#without doublets
clusters_to_keep <- level_01@meta.data %>% 
  group_by(., RNA_snn_res.0.4, is_endothlial,.drop =F) %>% 
  summarise(., count = n()) %>%
  group_by(RNA_snn_res.0.4) %>%
  mutate(., RNA_snn_res.0.4_cluster_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
  filter(., RNA_snn_res.0.4_cluster_purity > 0.95) %>% 
  pull(RNA_snn_res.0.4) %>% unique()

Idents(level_01) <- "RNA_snn_res.0.4"
level_02 <- subset(level_01, idents = clusters_to_keep)

```

Clean up list, remove Level 01

```{r}
rm(level_01); gc()
```

Clean up some metadata

```{r}
level_02$endothelial_UCell <- level_02$scGate_multi <- level_02$is_endothlial <- NULL

```

## Re-run single-cell workflow: harmony

Not using sctransform, regressing out %mito. Same function we used for Level 01

```{r}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #this is technically unneccessary, but also won't hurt
    FindVariableFeatures() %>% 
    ScaleData(., split.by = "batch") %>% 
    RunPCA() %>% 
    IntegrateLayers(object = ., method = HarmonyIntegration, orig.reduction = "pca",
  new.reduction = 'harmony', verbose = FALSE) %>% 
    RunUMAP(., dims = 1:30, reduction = "harmony") %>% #defaults
    FindNeighbors(., dims = 1:30, reduction = "harmony") %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1), reduction = "harmony")
}
```


```{r}
level_02 <- run_workflow(level_02)
```


## PCA correlations

```{r}
run_harmony_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$harmony@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, harmony_1, harmony_2, harmony_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "harmony_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "harmony_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_2") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "harmony_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_3") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```

```{r, fig.align='center', fig.height=5, fig.width=15}
wrap_plots(run_harmony_correlations(level_02), ncol = 3)
```

Correlation with harmony1

## UMAP overviews

Make UMAPs for features, ncounts, clusters and batch

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) 
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA"))
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) 
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("batch"), shuffle = T) 
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=12}
wrap_plots(plot_umap_dataset(level_02), ncol = 2)
```

## Clustree + Composition


## Clustree

```{r, fig.align='center', fig.height=10, fig.width=10}
clustree(level_02)
```

Stacked barplot for different resolutions capturing genotype split.

```{r, fig.align='center', fig.height=8, fig.width=25}
level_02@meta.data %>% 
    select(., batch, starts_with("RNA")) %>% 
    pivot_longer(., cols = starts_with("RNA"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("RNA_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, batch) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= batch)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =2) + 
    strip_fix 
```

## Feature UMAPs


```{r}
level_02_dotplot <- c(
  "prox1a", #LEC
  "cdh6", #LEC
  "stab2", #LEC, VEC
  "mrc1a",#LEC, VEC
  "lyve1b", #LEC, VEC
  # "cdh5", #AEC, VEC -> almost everything expresses this
  "flt1", #AEC
  "dll4", #AEC
  "hlx1", #migrAEC
  "esm1", #migrAEC
  "mki67", #cycling
  "pcna", #cycling
  "hand2", #endocardium
  "fn1a", #endocardium
  "osr2", #muLEC
  "sparc",
  "ifi30"
)
```


```{r}
run_feature_umap <- function(seurat, genes){
  FeaturePlot(seurat, features = genes, ncol = 4) 
}

```

```{r, fig.align='center', fig.height=20, fig.width=20}
run_feature_umap(seurat = level_02, genes = level_02_dotplot)
```


Make UMAP for res 0.4

```{r, fig.align='center', fig.height=5, fig.width=7}
 DimPlot(level_02, group.by = "RNA_snn_res.0.4", shuffle = T, label = T)
```

Show prox1a expression in big next to it?

```{r, fig.align='center', fig.height=5, fig.width=12}
 DimPlot(level_02, group.by = "RNA_snn_res.0.4", shuffle = T, label = T) +
  FeaturePlot(level_02, features = "prox1a")
```


## scGate: Find major subpopulations

Didn't really work, so we're skipping it. 

## Dotplot for genes

Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(sprintf("%s, %s",group, seurat@misc$name))  +plot_sizing_seurat + rotate_x
}

```

Make for res 0.4.

```{r, fig.align='center', fig.height=5, fig.width=7}
run_dotplot(seurat = level_02, group = "RNA_snn_res.0.4", genes = level_02_dotplot) + ylab('') + xlab("")
```


## Add annotations

### Cell type level

```{r}
level_02$L2_celltype <- case_when(
  level_02$RNA_snn_res.0.4 %in% c(1,12) ~ "VEC",
  level_02$RNA_snn_res.0.4 == 4 ~ "hippo+_VEC",
  level_02$RNA_snn_res.0.4 %in% c(6,8,9,10,11,14) ~ "AEC",
  level_02$RNA_snn_res.0.4 == 2 ~ "cycling_LEC_VEC",
  level_02$RNA_snn_res.0.4 == 3 ~ "cycling_AEC",
  level_02$RNA_snn_res.0.4 == 0 ~ "mixed_EC",
  level_02$RNA_snn_res.0.4 %in% c(5, 15) ~ "Endocardium",
  level_02$RNA_snn_res.0.4 == 7 ~ "LEC",
  level_02$RNA_snn_res.0.5 == 0 ~ "mixed_EC",
  level_02$RNA_snn_res.0.5 == 13 ~ "pre_muLEC"
)

```


### For each cluster

```{r}
add_seurat_names <- function(seurat, group, names){
  seurat 		<- SetIdent(seurat, value=group)
names(names) 		<- levels(seurat)
seurat 		<- RenameIdents(seurat, names)
seurat 		<- AddMetaData(
  object=seurat,
  metadata=seurat@active.ident,
  col.name="L2_seurat_celltype")
return(seurat)
  
}
```


```{r}

L2_seurat_names <- c(
'mixed_EC_01', #0
'VEC_01',#1
'cycling_LEC_VEC_01', #2
'cycling_AEC_01',#3
'hippo+_VEC_01',#4
'Endocardium_01',#5
'AEC_01', #6
'LEC_01', #7
'AEC_02', #8
'AEC_03',#9
'AEC_04', #10
"AEC_05", #11
"VEC_02", #12
"pre_muLEC_01", #13
"AEC_06", #14
"Endocardium_02" #15
)

level_02 <- add_seurat_names(
  seurat = level_02,
  group = "RNA_snn_res.0.4",
  names = L2_seurat_names)

```

Show on UMAPs

```{r}
run_celltype_umaps <-function(seurat){
  colours <- Polychrome::dark.colors(length(unique(seurat$L2_celltype))) %>% unname()
  DimPlot(seurat, group.by = "L2_celltype", cols = colours) + ggtitle(seurat@misc$name)
}

```

```{r,fig.align='center', fig.height=5, fig.width=7}
run_celltype_umaps(level_02)
```

## Export 

Export Level 02. 

```{r}

file_name <- here('output/data/Datsets/merged_wildtypes/merged_wildtypes_Level_02_annotated.qs2')
qs_save(level_02, file_name)

```



