---
title: "D10025 yap1 dataset: Make Level 02 (all endothelial cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 02 object for yap1 dataset. Alternative to ./pre_processing_D10025_yap1_dataset_makeObject_Level02_annotation.Rmd, but using SCT and regressing out cell cycle instead

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_D10025_yap1_dataset_makeObject_Level01_SCT_annotation.Rmd

```{r}
input_paths <- c(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_SCT_kept_doublets_annotated.qs2'),
                 here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_SCT_annotated.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: Endothelial clusters only

Subset Level 01 to only include endothelial subsets. By going through ./pre_processing_D10025_yap1_dataset_makeObject_Level01_SCT_annotation.Rmd, we can see that:

- there are no low quality clusters that stand out

- we can find endothelial subsets using a mix of marker genes and scGate endothelial purity (95%) for clusters


```{r}
#for each setting separately

#with doublets
cells_to_keep <- list_objects$D10025_yap1_dataset_Level_01_SCT_kept_doublets_annotated@meta.data %>% 
  filter(., SCT_snn_res.0.4_cluster_purity > 0.95) %>% 
  rownames_to_column("Barcode") %>% 
  pull(Barcode) 

list_objects$D10025_yap1_dataset_Level_02_kept_doublets <- subset(list_objects$D10025_yap1_dataset_Level_01_SCT_kept_doublets_annotated, cells = cells_to_keep)

#without doublets
cells_to_keep <- list_objects$D10025_yap1_dataset_Level_01_SCT_annotated@meta.data %>% 
  filter(., SCT_snn_res.0.4_cluster_purity > 0.95) %>% 
  rownames_to_column("Barcode") %>% 
  pull(Barcode) 

list_objects$D10025_yap1_dataset_Level_02 <- subset(list_objects$D10025_yap1_dataset_Level_01_SCT_annotated, cells = cells_to_keep)


```

Clean up list, remove Level 01

```{r}
list_objects$D10025_yap1_dataset_Level_01_SCT_kept_doublets_annotated <- list_objects$D10025_yap1_dataset_Level_01_SCT_annotated <- NULL
gc()
```

Clean up some metadata

```{r}
list_objects$D10025_yap1_dataset_Level_02_kept_doublets$endothelial_UCell <- list_objects$D10025_yap1_dataset_Level_02_kept_doublets$scGate_multi <- list_objects$D10025_yap1_dataset_Level_02_kept_doublets$is_endothlial <- list_objects$D10025_yap1_dataset_Level_02_kept_doublets$SCT_snn_res.0.4_cluster_purity <- NULL

list_objects$D10025_yap1_dataset_Level_02$endothelial_UCell <- list_objects$D10025_yap1_dataset_Level_02$scGate_multi <- list_objects$D10025_yap1_dataset_Level_02$is_endothlial <- list_objects$D10025_yap1_dataset_Level_02$SCT_snn_res.0.4_cluster_purity <- NULL
```


Replace name in object

```{r}
list_objects$D10025_yap1_dataset_Level_02_kept_doublets@misc$name <- "Level_02_kept_doublets"
list_objects$D10025_yap1_dataset_Level_02@misc$name <- "Level_02"

```



## Re-run single-cell workflow

Using SCTransform, regressing out %mito and tricycle position

```{r}
options(future.globals.maxSize = 30000 * 1024^2)
run_workflow <- function(seurat){
  seurat <- SCTransform(seurat, vars.to.regress = c("percent.mt", "tricycle_position")) %>%  #using glmGamPoi 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1))
  
  #add normalisation to RNA as well
  DefaultAssay(object = seurat) <- "RNA"
  seurat <- NormalizeData(seurat) %>%   #should deal with basic library size differences
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt","tricycle_position"))
  
  return(seurat)
}
```


```{r, message = FALSE}
list_objects <- lapply(list_objects, run_workflow)
```


## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "top with doublets, bottom without")
```


## UMAP overviews

Make UMAPs for features, ncounts, clusters, genotype and cell cycle.

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[5]] <- DimPlot(seurat,  group.by = c("tricycle_stage"), shuffle = T) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=25}
plot_list <- lapply(list_objects, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F)) + plot_annotation(title = "top with doublets, bottom without") + plot_layout(nrow = 2)
```

I'm not sure the regressing out of cell cycle is really doing anything.

## Clustree + Composition


```{r, fig.align='center', fig.height=12, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  clustree(seurat, prefix = c("SCT_snn_res.")) + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2)
```

Maybe res 0.4?

Stacked barplot for different resolutions capturing genotype split.

```{r}
run_barplot_genotype <- function(seurat){
  seurat@meta.data %>% 
    select(., Genotype, starts_with("SCT")) %>% 
    pivot_longer(., cols = starts_with("SCT"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("SCT_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, Genotype) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= Genotype)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =2) + 
    strip_fix +
    ggtitle(seurat@misc$name)
}
```


```{r, fig.align='center', fig.height=20, fig.width=30}

plot_list <- lapply(list_objects, run_barplot_genotype)

wrap_plots(plot_list, ncol = 1)
```

Mutant specific cluster jumps out from resolution 0.2.

Make UMAPs for res 0.3 and res 0.4

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- DimPlot(seurat, group.by = "SCT_snn_res.0.3", shuffle = T, label = T) + ggtitle(sprintf("res 0.3, %s", seurat@misc$name)) 
  plot_list[[2]] <- DimPlot(seurat, group.by = "SCT_snn_res.0.4", shuffle = T, label = T) + ggtitle(sprintf("res 0.4, %s", seurat@misc$name)) 
  return(plot_list)
})

wrap_plots(unlist(plot_list, recursive = F), ncol = 2)
```

## scGate: Find major subpopulations

Hasn't really been working, so we're skipping this.

## Feature UMAPs

```{r}
level_02_dotplot <- c(
  "prox1a", #LEC
  "cdh6", #LEC
  "stab2", #LEC, VEC
  "mrc1a",#LEC, VEC
  "lyve1b", #LEC, VEC
  # "cdh5", #AEC, VEC -> almost everything expresses this
  "flt1", #AEC
  "dll4", #AEC
  "hlx1", #migrAEC
  "esm1", #migrAEC
  "mki67", #cycling
  "pcna", #cycling
  "hand2", #endocardium
  "fn1a" #endocardium
)
```


This can sometimes be helpful.

```{r}
run_feature_umap <- function(seurat, genes){
  FeaturePlot(seurat, features = genes, ncol = 4) 
}

```

```{r, fig.align='center', fig.height=40, fig.width=20}
plot_list <- lapply(list_objects, function(seurat) run_feature_umap(seurat, level_02_dotplot))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```


## Dotplot for genes


Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(sprintf("%s, %s",group, seurat@misc$name))  +plot_sizing_seurat + rotate_x
}

```

Make for res 0.4.

```{r, fig.align='center', fig.height=8, fig.width=10}
plot_list <- lapply(list_objects, function(seurat) run_dotplot(seurat, level_02_dotplot, group = "SCT_snn_res.0.4"))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```


## Add annotations

### Cell type level

```{r}
#For with doublets
list_objects$D10025_yap1_dataset_Level_02_kept_doublets$L2_celltype <- case_when(
  list_objects$D10025_yap1_dataset_Level_02_kept_doublets$SCT_snn_res.0.4 %in% c(0,10,13,4,5) ~ "VEC",
  list_objects$D10025_yap1_dataset_Level_02_kept_doublets$SCT_snn_res.0.4 %in% c(11,3,6,7,8,9) ~ "AEC",
  list_objects$D10025_yap1_dataset_Level_02_kept_doublets$SCT_snn_res.0.4 == 1 ~ "cycling_EC",
  list_objects$D10025_yap1_dataset_Level_02_kept_doublets$SCT_snn_res.0.4 %in% c(2,14) ~ "Endocardium",
  list_objects$D10025_yap1_dataset_Level_02_kept_doublets$SCT_snn_res.0.4 == 12 ~ "LEC"
)

#for without doublets
list_objects$D10025_yap1_dataset_Level_02$L2_celltype <- case_when(
  list_objects$D10025_yap1_dataset_Level_02$SCT_snn_res.0.4 %in% c(0,2,5,10,13) ~ "VEC",
  list_objects$D10025_yap1_dataset_Level_02$SCT_snn_res.0.4 %in% c(4,6,7,8,9,12) ~ "AEC",
  list_objects$D10025_yap1_dataset_Level_02$SCT_snn_res.0.4 ==1 ~ "cycling_EC",
  list_objects$D10025_yap1_dataset_Level_02$SCT_snn_res.0.4 ==3 ~ "Endocardium",
  list_objects$D10025_yap1_dataset_Level_02$SCT_snn_res.0.4 == 11 ~ "LEC"
)

```


### For each cluster

```{r}
add_seurat_names <- function(seurat, group, names){
  seurat 		<- SetIdent(seurat, value=group)
names(names) 		<- levels(seurat)
seurat 		<- RenameIdents(seurat, names)
seurat 		<- AddMetaData(
  object=seurat,
  metadata=seurat@active.ident,
  col.name="L2_seurat_celltype")
return(seurat)
  
}
```


```{r}
#for with doublets
L2_seurat_names <- c(
'VEC_01', #0
'cycling_EC_01',#1
'Endocardium_01', #2
'AEC_01',#3
'VEC_02',#4
'VEC_03',#5
'AEC_02', #6
'AEC_03', #7
'AEC_04', #8
'AEC_05',#9
'VEC_04', #10
"AEC_06", #11
"LEC_01", #12
"VEC_05" ,#13
"Endocardium_02" #14
)

list_objects$D10025_yap1_dataset_Level_02_kept_doublets <- add_seurat_names(
  seurat = list_objects$D10025_yap1_dataset_Level_02_kept_doublets,
  group = "SCT_snn_res.0.4",
  names = L2_seurat_names)

#for without doublets
L2_seurat_names <- c(
'VEC_01', #0
'cycling_EC_01',#1
'VEC_02', #2
'Endocardium_01',#3
'AEC_01',#4
'VEC_03',#5
'AEC_02', #6
'AEC_03', #7
'AEC_04', #8
'AEC_05',#9
'VEC_04', #10
"LEC_01", #11
"AEC_06", #12
"VEC_05" #13
)

list_objects$D10025_yap1_dataset_Level_02 <- add_seurat_names(
  seurat = list_objects$D10025_yap1_dataset_Level_02,
  group = "SCT_snn_res.0.4",
  names = L2_seurat_names)

```

Show on UMAPs

```{r}
run_celltype_umaps <-function(seurat){
  colours <- Polychrome::dark.colors(5); names(colours) <- c("AEC", "VEC", "LEC", 
                                                             "cycling_EC", "Endocardium")
  DimPlot(seurat, group.by = "L2_celltype", cols = colours) + ggtitle(seurat@misc$name)
}

```

```{r,fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, run_celltype_umaps)

wrap_plots(plot_list, ncol =2) + plot_layout(guides = "collect")

```

## Export 

Export Level 02. 

```{r}

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_SCT_annotated.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_02, file_name)

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_SCT_kept_doublets_annotated.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_02_kept_doublets , file_name)

```



