---
title: "D10051 meox1 dataset: Annotate Level 01 (all cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Annotate Level 01 of the meox1 dataset


### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(tricycle)
  library(scGate)
})
```

### Load objects

Objects made here: ./pre_processing_D10051_meox1_dataset_makeObject_Level01.Rmd 

Only loading merged ones for now.

```{r}
input_paths <- c(here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_01_kept_doublets.qs2'),
                here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_01.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

Give them names in the object

```{r}
list_objects$D10051_meox1_dataset_Level_01_kept_doublets@misc[["name"]] <- "Level_01_kept_doublets"
list_objects$D10051_meox1_dataset_Level_01@misc[["name"]] <- "Level_01"

```


Read in mouse -> fish names

```{r}
mouse_to_fish <- qs_read(here("data/genesets/geneMap_mmusculus_to_drerio.qs2"))
```


### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```

## Tricycle: Run cell cycle analysis

Make function.

```{r}

run_tricycle_for_fish <- function(seurat, x_species_dictionary = mouse_to_fish){
  #prepare
  sce <- as.SingleCellExperiment(seurat, assay = "RNA") #convert to SCE
  sub_sce <- sce[rownames(sce) %in% x_species_dictionary$drerio_zfin_id_symbol,] #subset to genes in dic
  sub_x_species_dictionary <- x_species_dictionary %>%  #subset dictionary to only include the genes that we kept 
  filter(., drerio_zfin_id_symbol %in% rownames(sub_sce)) %>% 
  distinct(drerio_zfin_id_symbol, .keep_all = TRUE) #needed for function
  sub_x_species_dictionary <- sub_x_species_dictionary[match(rownames(sub_sce),
                                                             sub_x_species_dictionary$drerio_zfin_id_symbol),]
  stopifnot(all(sub_x_species_dictionary$drerio_zfin_id_symbol == rownames(sub_sce))) #check
  
  #run tricycle
  sub_sce <- project_cycle_space(sub_sce, 
                                 gname = sub_x_species_dictionary$mmusculus_mgi_symbol, 
                                 species = "mouse", 
                                 gname.type = "SYMBOL")
  #estimate position
  sub_sce <- estimate_cycle_position(sub_sce)
  
  
  #add back to seurat object
  seurat$tricycle_position <- sub_sce$tricyclePosition
  
  #add categorical cell cycle:
  # https://github.com/lb15/multiciliation_cycle/blob/main/mcc_timecourse/mcc_timecourse_analysis.R#L619
  seurat$tricycle_stage <- case_when(seurat$tricycle_position > pi/2 & seurat$tricycle_position < pi ~ "S",
                             seurat$tricycle_position >= pi & seurat$tricycle_position < pi*7/4 ~ "G2M",
                             TRUE ~ "G1/G0")
  
  rm(sce,sub_sce, sub_x_species_dictionary); gc()
  return(seurat)
  

}

```


```{r}
list_objects <- lapply(list_objects, run_tricycle_for_fish)
```

### Visualise test genes

Write function:

```{r}

make_tricycle_plots <- function(seurat){
  #colours 
  cols_cellcycle <- c('#ffff66', '#a0d0e0', '#cc85ff')
  names(cols_cellcycle) <- c('G1', 'G2M', 'S')
  rects <- data.frame(xmin = c(0, 0.5*pi, pi, 1.75*pi),
                    xmax = c(0.5*pi, pi, 1.75*pi, 2*pi),
                    fill = c('#ffff66',  '#cc85ff','#a0d0e0', '#ffff66'))
  
  
  plot_list <- list()
  plot_list[[1]] <- fit_periodic_loess(seurat$tricycle_position,
                                       seurat[["RNA"]]@layers$data[which(rownames(seurat) == 'cdt1'),],
                                       plot = TRUE,
                                       x_lab = "Cell cycle position \u03b8", y_lab = "cdt1 expression",
                                       fig.title = paste0("Expression of cdt1 (G1-phase)"))$fig +
    plot_sizing +
    annotate(geom = "text", x = c(0.25*pi, 0.75*pi, 1.4*pi, 1.9*pi), y = 4, label = c("G0/G1", "S", "G2M","G0/G1"), size = 4) +
    coord_cartesian(expand = FALSE, clip = "off", ylim = c(0, 4.5)) +
    geom_vline(xintercept=0.5*pi,linetype="dashed")+
    geom_vline(xintercept =pi,linetype="dashed")+
    geom_vline(xintercept=1.75*pi,linetype="dashed")  +
    geom_rect(aes(xmin = xmin, xmax = xmax, fill = fill, ymin = -Inf, ymax = Inf),
              data = rects, inherit.aes = FALSE, size = 1, alpha = .1) +
    scale_fill_manual(values = c( '#a0d0e0', '#cc85ff', '#ffff66','#ffff66')) +
    theme(legend.position = "none")


  plot_list[[2]] <- fit_periodic_loess(seurat$tricycle_position,
                                       seurat[["RNA"]]@layers$data[which(rownames(seurat) == 'pcna'),],
                                       plot = TRUE,
                                       x_lab = "Cell cycle position \u03b8", y_lab = "pcna expression",
                                       fig.title = paste0("Expression of pcna (S-phase)"))$fig + 
    plot_sizing +
    annotate(geom = "text", x = c(0.25*pi, 0.75*pi, 1.4*pi, 1.9*pi), y = 4, label = c("G0/G1", "S", "G2M","G0/G1"), size = 4) +
    coord_cartesian(expand = FALSE, clip = "off", ylim = c(0, 6)) +
    geom_vline(xintercept=0.5*pi,linetype="dashed")+
    geom_vline(xintercept =pi,linetype="dashed")+
    geom_vline(xintercept=1.75*pi,linetype="dashed") +
    geom_rect(aes(xmin = xmin, xmax = xmax, fill = fill, ymin = -Inf, ymax = Inf),
              data = rects, inherit.aes = FALSE, size = 1, alpha = .1) +
    scale_fill_manual(values = c( '#a0d0e0', '#cc85ff', '#ffff66','#ffff66')) +
    theme(legend.position = "none")
  
  plot_list[[3]] <- fit_periodic_loess(seurat$tricycle_position,
                                       seurat[["RNA"]]@layers$data[which(rownames(seurat) == 'mki67'),],
                                       plot = TRUE,
                                       x_lab = "Cell cycle position \u03b8", y_lab = "mki67 expression",
                                       fig.title = paste0("Expression of mki67 (G2M-phase)"))$fig +
    plot_sizing +
    annotate(geom = "text", x = c(0.25*pi, 0.75*pi, 1.4*pi, 1.9*pi), y = 4, label = c("G0/G1", "S", "G2M","G0/G1"), size = 4) +
    coord_cartesian(expand = FALSE, clip = "off", ylim = c(0, 4.5)) +
    geom_vline(xintercept=0.5*pi,linetype="dashed")+
    geom_vline(xintercept =pi,linetype="dashed")+
    geom_vline(xintercept=1.75*pi,linetype="dashed") +
    geom_rect(aes(xmin = xmin, xmax = xmax, fill = fill, ymin = -Inf, ymax = Inf),
              data = rects, inherit.aes = FALSE, size = 1, alpha = .1) +
    scale_fill_manual(values = c( '#a0d0e0', '#cc85ff', '#ffff66','#ffff66')) +
    theme(legend.position = "none")
  
  return(plot_list)
}


```

Run for all

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, make_tricycle_plots)

wrap_plots(unlist(plot_list, recursive = F)) + plot_annotation(title = "top = with doublets, bottom = without")
```

### UMAPs for cell cycle stages

```{r}
run_umap_stages <- function(seurat){
  #colours 
  cols_cellcycle <- c('#ffff66', '#a0d0e0', '#cc85ff')
  names(cols_cellcycle) <- c('G1/G0', 'G2M', 'S')
  #plot
  DimPlot(seurat, group.by = "tricycle_stage", cols = cols_cellcycle) + ggtitle(seurat@misc$name)
}
```

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, run_umap_stages)

wrap_plots(plot_list, ncol = 2) + plot_layout(guides = "collect", axis_titles = "collect")
```



## Clustree

```{r, fig.align='center', fig.height=12, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  clustree(seurat) + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2)
```

Maybe try 0.5? Might already be too granular. 

## Low quality clusters

Do any clusters have high mt and/or low count? Being mindful that some celltypes do have naturally higher %mt and counts.

```{r, fig.align='center', fig.height=8, fig.width=15}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = RNA_snn_res.0.5, fill=RNA_snn_res.0.5)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[2]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nFeature_RNA), x = RNA_snn_res.0.5, fill=RNA_snn_res.0.5)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[3]] <- seurat@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = RNA_snn_res.0.5, fill=RNA_snn_res.0.5)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  return(plot_list) 
})

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

There are definitively a few clusters with smaller library size, though they don't necessarily seem to express many mito genes. 

Make DimPlots for this resolution.

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, function(seurat){
  DimPlot(seurat, group.by = "RNA_snn_res.0.5", shuffle = T, label = T)
})

wrap_plots(plot_list, ncol = 2)
```

I really hate this, nmads doesn't work, and its very difficult to say whether these cells are bad quality or not- because in terms of number of genes they do still express quite a decent amount. The two peaks are just not right. 

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()

  #ncount
  plot_list[[1]] <- ggplot(seurat@meta.data, aes(x=log10(nCount_RNA))) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('log10(count)') +
    plot_sizing + ggtitle(sprintf("Library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggplot(seurat@meta.data, aes(x=log10(nFeature_RNA))) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('log10(nfeature)') +
    plot_sizing + ggtitle(sprintf("#features, %s",  seurat@project.name))
  
  
  return(plot_list)
  
})

wrap_plots(unlist(plot_list, recursive = F), ncol =2)
```

It particularly looks like the features were not affected by the ncount threshold of 600 at all (this is roughly what the yap1 dataset had)

```{r}
list_objects$D10051_meox1_dataset_Level_01_kept_doublets$nFeature_RNA %>% min()
```

Yeah, that's too low. We need to implement another round of filtering. Feature threshold in the yap1 dataset was ~400. Let's use this and re-run.

## Additional QC filtering step

Try some more stringent feature thresholds

```{r}
remove_outliers <- function(seurat){
  countfeatures_low <- seurat$nFeature_RNA < 400 
  seurat <- seurat[, !countfeatures_low]
}
```

```{r}
list_objects <- lapply(list_objects, remove_outliers)
```


### replot distribution of counts and features

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()

  #ncount
  plot_list[[1]] <- ggplot(seurat@meta.data, aes(x=log10(nCount_RNA))) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('log10(count)') +
    plot_sizing + ggtitle(sprintf("Library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggplot(seurat@meta.data, aes(x=log10(nFeature_RNA))) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('log10(nfeature)') +
    plot_sizing + ggtitle(sprintf("#features, %s",  seurat@project.name))
  
  
  return(plot_list)
  
})

wrap_plots(unlist(plot_list, recursive = F), ncol =2)
```

## Run single-cell workflow

Not using sctransform, regressing out %mito

```{r}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #should deal with basic library size differences
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1))
}
```

```{r, message=FALSE}
list_objects <- lapply(list_objects, run_workflow)
```

### PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F)) + plot_annotation("top with doublets, bottom without")
```

### UMAP overviews

Not that much to see, only make UMAPs for nfeature, ncount and seurat_cluster.

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=20}
plot_list <- lapply(list_objects, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F), ncol = 4) 
```

### Stacked barplots for Genotype

```{r}
run_stacked_barplot <- function(seurat){
  seurat@meta.data %>% 
    select(., starts_with("RNA"), Genotype) %>% 
    pivot_longer(., cols = starts_with("RNA"), names_to = "resolution", values_to = "cluster") %>% 
    group_by(resolution, cluster, Genotype) %>% 
    summarise(., count =n()) %>% 
    ggplot(., aes(x = cluster, y = count, fill = Genotype))+
    geom_bar(position = "fill", stat = "identity") +
    facet_wrap(~resolution, scales = "free_x", ncol =5) + plot_sizing+ strip_fix

}
```


```{r, fig.align='center', fig.height=15, fig.width=30}
plot_list <- lapply(list_objects, run_stacked_barplot)

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- DimPlot(seurat, group.by = "RNA_snn_res.0.1", shuffle = T, label = T)
  plot_list[[2]] <- FeaturePlot(seurat, features  = c("fn1a", "lyve1b", "prox1a"), ncol = 3)
  return(plot_list)
})

wrap_plots(unlist(plot_list, recursive = F), ncol = 2) +plot_layout(widths = c(1,3))
```

Can already spot the hippo+ cluster, and it stands out with a wt enrichment as well! promising.

### Low quality clusters

Re-run this

```{r, fig.align='center', fig.height=8, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = RNA_snn_res.0.5, fill=RNA_snn_res.0.5)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[2]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nFeature_RNA), x = RNA_snn_res.0.5, fill=RNA_snn_res.0.5)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[3]] <- seurat@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = RNA_snn_res.0.5, fill=RNA_snn_res.0.5)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  return(plot_list) 
})

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

Definitively still has some clusters that are lower... 

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, function(seurat){
  DimPlot(seurat, group.by = "RNA_snn_res.0.5", shuffle = T, label = T)
})

wrap_plots(plot_list, ncol = 2)
```

Is there a sample difference?

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, function(seurat){
seurat@meta.data %>% 
  ggplot(., aes(y = log10(nFeature_RNA), x = RNA_snn_res.0.5, fill=Genotype)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
})

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

Not really. 

### For kept doublets - where are doublets?

```{r, fig.align='center', fig.height=5, fig.width=7}
DimPlot(list_objects$D10051_meox1_dataset_Level_01_kept_doublets, group.by = "scDblFinder.class", shuffle = T)
```

There's defintively quite a few in the endothelial subset, and something weird's going on in one of the clusters.

### Comment on having an even more stringent filter

At this stage I can't really justify removing more cells by increasing the thresholds, though the data distribution does really look strange. 


## scGate: Find endothelial subset

Using scGate as a computational gating strategy to identify all endothelial cells. Endothelial cells are marked by expression of kdrl, and absence of expression of col1a1a (Fibroblasts), epcam (Epithelial), gata2b (haematopoietic), myod1 (Muscle). 

Mark clusters that have > 95% purity for resolution 0.5.

```{r}
endothelial_gating_model <- gating_model(name = "endothelial", signature = c("kdrl+", 
                                                                    "col1a1a-",
                                                                    "epcam-",
                                                                    "gata2b-", 
                                                                    "myod1-", 
                                                                    "gch2-",
                                                                    "aox5-"))

#write function
run_scgate_endo <- function(seurat){
  seurat <- scGate(data = seurat,
                   model = endothelial_gating_model,
                   output.col.name = "is_endothlial")

  return(seurat)
}

```


```{r}
list_objects <- lapply(list_objects, run_scgate_endo)
```

Show UMAPs with is pure

```{r}
run_umap_scgate_results <- function(seurat){
  #colours
  pure_colours <- c("grey", "green3"); names(pure_colours) <- c("Impure", "Pure")
  DimPlot(seurat, group.by = "is_endothlial", cols = pure_colours) + ggtitle(sprintf("Endo Gate, %s", seurat@misc$name))
}
```

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, run_umap_scgate_results)

wrap_plots(plot_list, ncol = 2) + plot_layout(guides = "collect", axis_titles = "collect")
```

Make barplots for endo purity

```{r, fig.align='center', fig.height=5, fig.width=12}
plot_list <- lapply(list_objects, function(seurat){
  seurat@meta.data %>% 
  mutate(.,RNA_snn_res.0.5 = as.factor(RNA_snn_res.0.5)) %>% 
  group_by(., RNA_snn_res.0.5, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.5) %>%
    mutate(., RNA_snn_res.0.5_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.5_purity) %>% distinct() %>% 
    ggplot(., aes(x= reorder(RNA_snn_res.0.5, -RNA_snn_res.0.5_purity),  y = RNA_snn_res.0.5_purity))+
    geom_bar(stat = "identity") + plot_sizing +
    ylab("Endo Purity per cluster") + xlab("cluster") + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2) + plot_layout(guides = "collect", axis_titles = "collect", axes = "collect_y")

```

## Dotplot for genes

Use some known marker genes for celltypes to double check scGate annotation.

```{r}
dotplot.level.01 <- c(
  "cdh6",
  "prox1a",
  "stab2",
  "lyve1b",
  "mrc1a",
  "flt1",
  "dll4",
  "esm1",
  "hlx1",
  "hand2",
  "fn1a",
  "gata2b",
  "runx1",
  "alas2",
  "blvrb",
  "spi1b",
  "mpx",
  "neurod4",
  "nova2",
  "cdh1",
  "epcam",
  "myod1",
  "myog",
  "pmela",
  "dct",
  "lum",
  "col1a1a",
  "krt4",
  "cyt1",
  "cryba1b",
  "crybb1",
  "ednrba",
  "pnp4a",
  "gch2",
  "aox5")
```

Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(seurat@misc$name)  +plot_sizing_seurat + rotate_x
}

```

Make dotplots for all clusters

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, function(seurat) run_dotplot(seurat, dotplot.level.01, group = "RNA_snn_res.0.5"))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

Make dotplot for clusters that have mid-range purity.

```{r}
run_dotplot_mid_purity <- function(seurat, genes, limits){
  clusters <- seurat@meta.data %>% 
  mutate(.,RNA_snn_res.0.5 = as.factor(RNA_snn_res.0.5)) %>% 
  group_by(., RNA_snn_res.0.5, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.5) %>%
    mutate(., RNA_snn_res.0.5_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.5_purity) %>%
    filter(., RNA_snn_res.0.5_purity > limits[1] &  RNA_snn_res.0.5_purity < limits[2] ) %>% 
    pull(RNA_snn_res.0.5)
  
  seurat <- SetIdent(seurat, value = "RNA_snn_res.0.5")
  DotPlot(seurat, features = genes, group.by = "RNA_snn_res.0.5", idents = clusters) + ggtitle(seurat@misc$name)  +plot_sizing_seurat + rotate_x
}

```

```{r, fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, function(seurat) run_dotplot_mid_purity(seurat, dotplot.level.01, limits = c(0.4, 0.95)))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

In kept doublet setting, I'd say it looks fine. 

Looking more into cluster 10 in the doublet removed setting. They're prox1a+, so they're probably LECs. What's the purity there?

```{r}
list_objects$D10051_meox1_dataset_Level_01@meta.data %>% 
  mutate(.,RNA_snn_res.0.5 = as.factor(RNA_snn_res.0.5)) %>% 
  group_by(., RNA_snn_res.0.5, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.5) %>%
    mutate(., RNA_snn_res.0.5_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.5_purity) %>%
    filter(., RNA_snn_res.0.5_purity > 0.4 &  RNA_snn_res.0.5_purity < 0.95) %>% 
  distinct()
```

Ah right, so 93% purity.. That's fair we'll include it! 


## Export 

Export them all. Selection of clusters to move to Level 02 will happen in next script

```{r}

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10051_meox1_dataset_Level_01_annotated.qs2')
qs_save(list_objects$D10051_meox1_dataset_Level_01, file_name)

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10051_meox1_dataset_Level_01_kept_doublets_annotated.qs2')
qs_save(list_objects$D10051_meox1_dataset_Level_01_kept_doublets , file_name)

```



