---
title: "D10051 meox1 dataset: Make Level 02 (all endothelial cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 02 object for meox1 dataset, using a more stringent filtering threshold to ditch all the low quality cells. 

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_D10051_meox1_dataset_makeObject_Level01_annotation_stringent.Rmd

```{r}
input_paths <- c(here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_01_stringent_kept_doublets_annotated.qs2'),
                 here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_01_stringent_annotated.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: Endothelial clusters only

Subset Level 01 to only include endothelial subsets. ./pre_processing_D10051_meox1_dataset_makeObject_Level01_annotation_stringent.Rmd, for why we pick 90% purity threshold on res 0.5.

```{r,message=FALSE}
#for each setting separately

#with doublets
clusters_to_keep <- list_objects$D10051_meox1_dataset_Level_01_stringent_kept_doublets_annotated@meta.data %>% 
  mutate(.,RNA_snn_res.0.5 = as.factor(RNA_snn_res.0.5)) %>% 
  group_by(., RNA_snn_res.0.5, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.5) %>%
    mutate(., RNA_snn_res.0.5_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.5_purity) %>% 
  filter(., RNA_snn_res.0.5_purity > 0.90) %>% 
  pull(RNA_snn_res.0.5) %>% unique()

cells_to_keep <- list_objects$D10051_meox1_dataset_Level_01_stringent_kept_doublets_annotated@meta.data %>% 
  rownames_to_column("Barcode") %>% 
  filter(., RNA_snn_res.0.5 %in% clusters_to_keep) %>% 
  pull(Barcode)

list_objects$D10051_meox1_dataset_Level_02_kept_doublets <- subset(list_objects$D10051_meox1_dataset_Level_01_stringent_kept_doublets_annotated, cells = cells_to_keep)

#without doublets
clusters_to_keep <- list_objects$D10051_meox1_dataset_Level_01_stringent_annotated@meta.data %>% 
 mutate(.,RNA_snn_res.0.5 = as.factor(RNA_snn_res.0.5)) %>% 
  group_by(., RNA_snn_res.0.5, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.5) %>%
    mutate(., RNA_snn_res.0.5_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.5_purity) %>% 
  filter(., RNA_snn_res.0.5_purity > 0.90) %>% 
  pull(RNA_snn_res.0.5) %>% unique()

cells_to_keep <- list_objects$D10051_meox1_dataset_Level_01_stringent_annotated@meta.data %>% 
  rownames_to_column("Barcode") %>% 
  filter(., RNA_snn_res.0.5 %in% clusters_to_keep) %>% 
  pull(Barcode)

list_objects$D10051_meox1_dataset_Level_02 <- subset(list_objects$D10051_meox1_dataset_Level_01_stringent_annotated, cells = cells_to_keep)


```

Clean up list, remove Level 01

```{r}
list_objects$D10051_meox1_dataset_Level_01_stringent_kept_doublets_annotated <- list_objects$D10051_meox1_dataset_Level_01_stringent_annotated <- NULL
gc()
```

Clean up some metadata

```{r}
list_objects$D10051_meox1_dataset_Level_02_kept_doublets$endothelial_UCell <- list_objects$D10051_meox1_dataset_Level_02_kept_doublets$scGate_multi <- list_objects$D10051_meox1_dataset_Level_02_kept_doublets$is_endothlial <- NULL

list_objects$D10051_meox1_dataset_Level_02$endothelial_UCell <- list_objects$D10051_meox1_dataset_Level_02$scGate_multi <- list_objects$D10051_meox1_dataset_Level_02$is_endothlial  <- NULL
```


Replace name in object

```{r}
list_objects$D10051_meox1_dataset_Level_02_kept_doublets@misc$name <- "Level_02_kept_doublets"
list_objects$D10051_meox1_dataset_Level_02@misc$name <- "Level_02"

```



## Re-run single-cell workflow

Not using sctransform, regressing out %mito. Same function we used for Level 01

```{r}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #this is technically unneccessary, but also won't hurt
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1)) #this will overwrite old clustering (that's okay)
}
```


```{r, message=FALSE}
list_objects <- lapply(list_objects, run_workflow)
```


## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F), ncol =3) +plot_annotation(title = "top with doublets, bottom without")
```


## UMAP overviews

Make UMAPs for features, ncounts, clusters, genotype and cell cycle.

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[5]] <- DimPlot(seurat,  group.by = c("tricycle_stage"), shuffle = T) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=25}
plot_list <- lapply(list_objects, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F)) + plot_annotation(title = "top with doublets, bottom without") + plot_layout(nrow = 2)
```

Is that a little mutant specific cluster??

## Clustree + Composition


```{r, fig.align='center', fig.height=12, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  clustree(seurat) + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2)
```

0.2 or 0.3?

Stacked barplot for different resolutions capturing genotype split.

```{r}
run_barplot_genotype <- function(seurat){
  seurat@meta.data %>% 
    select(., Genotype, starts_with("RNA")) %>% 
    pivot_longer(., cols = starts_with("RNA"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("RNA_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, Genotype) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= Genotype)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =2) + 
    strip_fix +
    ggtitle(seurat@misc$name)
}
```


```{r, fig.align='center', fig.height=20, fig.width=30}

plot_list <- lapply(list_objects, run_barplot_genotype)

wrap_plots(plot_list, ncol = 1)
```

0.3 starts capturing the mutant specific cluster. 

Make UMAPs for res 0.3

```{r, fig.align='center', fig.height=5, fig.width=12}
plot_list <- lapply(list_objects, function(seurat){
  DimPlot(seurat, group.by = "RNA_snn_res.0.3", shuffle = T, label = T) + ggtitle(sprintf("res 0.3, %s", seurat@misc$name)) 
})

wrap_plots(plot_list, ncol = 2)
```

Mutant-enriched cluster 11/12?

## Low quality clusters

Do any clusters have high mt and/or low count? Being mindful that some celltypes do have naturally higher %mt and counts.

```{r, fig.align='center', fig.height=8, fig.width=15}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = RNA_snn_res.0.3, fill=RNA_snn_res.0.3)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[2]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nFeature_RNA), x = RNA_snn_res.0.3, fill=RNA_snn_res.0.3)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[3]] <- seurat@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = RNA_snn_res.0.3, fill=RNA_snn_res.0.3)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  return(plot_list) 
})

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```


## Feature UMAPs

```{r}
level_02_dotplot <- c(
  "prox1a", #LEC
  "cdh6", #LEC
  "stab2", #LEC, VEC
  "mrc1a",#LEC, VEC
  "lyve1b", #LEC, VEC
  # "cdh5", #AEC, VEC -> almost everything expresses this
  "flt1", #AEC
  "dll4", #AEC
  "hlx1", #migrAEC
  "esm1", #migrAEC
  "mki67", #cycling
  "pcna", #cycling
  "hand2", #endocardium
  "fn1a", #endocardium
  "osr2", #muLEC
  "sparc"
)
```

This can sometimes be helpful.

```{r}
run_feature_umap <- function(seurat, genes){
  FeaturePlot(seurat, features = genes, ncol = 4) 
}

```

```{r, fig.align='center', fig.height=40, fig.width=20}
plot_list <- lapply(list_objects, function(seurat) run_feature_umap(seurat, level_02_dotplot))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

## Dotplot for genes


Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(sprintf("%s, %s",group, seurat@misc$name))  +plot_sizing_seurat + rotate_x
}

```

Make for res 0.3.

```{r, fig.align='center', fig.height=10, fig.width=10}
plot_list <- lapply(list_objects, function(seurat) run_dotplot(seurat, level_02_dotplot, group = "RNA_snn_res.0.3"))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

## Doublet UMAP

```{r, fig.align='center', fig.height=5, fig.width=7}
DimPlot(list_objects$D10051_meox1_dataset_Level_02_kept_doublets, group.by = "scDblFinder.class")
```

No obvious enrichments. 


## Add annotations

### Cell type level

```{r}
#For with doublets
list_objects$D10051_meox1_dataset_Level_02_kept_doublets$L2_celltype <- case_when(
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 %in% c(1,4) ~ "VEC",
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 %in% c(5,14) ~ "LEC",
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 %in% c(3,8,9,11,12,13) ~ "AEC",
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 == 0 ~ "cycling_EC",
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 %in% c(2,6) ~ "cycling_AEC",
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 %in% c(7, 15) ~ "Endocardium",
  list_objects$D10051_meox1_dataset_Level_02_kept_doublets$RNA_snn_res.0.3 == 10 ~ "pre_muLEC"
)

#for without doublets
list_objects$D10051_meox1_dataset_Level_02$L2_celltype <- case_when(
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 %in% c(1,4) ~ "VEC",
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 %in% c(6,13) ~ "LEC",
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 %in% c(2, 7,8,9,12) ~ "AEC",
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 == 0 ~ "cycling_EC",
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 %in% c(3,8,11) ~ "cycling_AEC",
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 == 5 ~ "Endocardium",
  list_objects$D10051_meox1_dataset_Level_02$RNA_snn_res.0.3 == 10 ~ "pre_muLEC"
)

```


### For each cluster

```{r}
add_seurat_names <- function(seurat, group, names){
  seurat 		<- SetIdent(seurat, value=group)
names(names) 		<- levels(seurat)
seurat 		<- RenameIdents(seurat, names)
seurat 		<- AddMetaData(
  object=seurat,
  metadata=seurat@active.ident,
  col.name="L2_seurat_celltype")
return(seurat)
  
}
```


```{r}
#for with doublets
L2_seurat_names <- c(
'cycling_EC_01', #0
'VEC_01',#1
'cycling_AEC_01', #2
'AEC_01',#3
'VEC_02',#4
'LEC_01',#5
'cycling_AEC_02', #6
'Endocardium_01', #7
'AEC_02', #8
'AEC_02',#9
'pre_muLEC_01', #10
'AEC_03', #11
'AEC_04', #12
'AEC_05', #13
'LEC_02', #14
'Endocardium_02' #15
)

list_objects$D10051_meox1_dataset_Level_02_kept_doublets <- add_seurat_names(
  seurat = list_objects$D10051_meox1_dataset_Level_02_kept_doublets,
  group = "RNA_snn_res.0.3",
  names = L2_seurat_names)

#for without doublets
L2_seurat_names <-  c(
'cycling_EC_01', #0
'VEC_01',#1
'AEC_01', #2
'cycling_AEC_01',#3
'VEC_02',#4
'Endocardium_01',#5
'LEC_01', #6
'AEC_02', #7
'cycling_AEC_02', #8
'AEC_03',#9
'pre_muLEC_01', #10
'cycling_AEC_03', #11
'AEC_04', #12
'LEC_02' #13
)

list_objects$D10051_meox1_dataset_Level_02 <- add_seurat_names(
  seurat = list_objects$D10051_meox1_dataset_Level_02,
  group = "RNA_snn_res.0.3",
  names = L2_seurat_names)

```

Show on UMAPs

```{r}
run_celltype_umaps <-function(seurat){
  colours <- Polychrome::dark.colors() %>% unname()
  DimPlot(seurat, group.by = "L2_celltype", cols = colours) + ggtitle(seurat@misc$name)
}

```

```{r,fig.align='center', fig.height=5, fig.width=10}
plot_list <- lapply(list_objects, run_celltype_umaps)

wrap_plots(plot_list, ncol =2) + plot_layout(guides = "collect")

```



## Export 

Export Level 02. 

```{r}

file_name <- here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_02_stringent_annotated.qs2')
qs_save(list_objects$D10051_meox1_dataset_Level_02, file_name)

file_name <- here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_02_stringent_kept_doublets_annotated.qs2')
qs_save(list_objects$D10051_meox1_dataset_Level_02_kept_doublets , file_name)

```



