---
title: "Combined WT only dataset: Make Level 01 (all cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Combine wildtype from yap1 and meox1 experiments. 


### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
 })
```

### Load data

From here ./pre_processing_D10025_yap1_dataset_makeObject_Level01.Rmd and here pre_processing_D10051_meox1_dataset_makeObject_Level01.Rmd. Note that each sample was processed there, e.g. QC was done on each sample separately.

Running with doublets removed: 

```{r}
yap1_wildtype <- qs_read(here("output/data/Samples/S20128/S20128_wildtype_Level_01.qs2"))
meox1_wildtype <- qs_read(here("output/data/Samples/S20201/S20201_wildtype_Level_01.qs2"))
```

Add batch information. 

```{r}
yap1_wildtype$batch <- "yap1_experiment"
meox1_wildtype$batch <- "meox1_experiment"
```


### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```



## Combine 

Try to just combine

```{r}
merged <- merge(yap1_wildtype,
                y = meox1_wildtype,
                project = "wildtype_only", merge.data = TRUE) 

merged<- JoinLayers(merged) #seurat v5 things

```


Single-cell workflow:

Regress out %mito and batch

```{r}
run_workflow_batch <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #should deal with basic library size differences
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt", "batch")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = 0.8)
  return(seurat)
    # FindClusters(., resolution = seq(0.1, 1, 0.1))
}

```

```{r}
merged <- run_workflow_batch(merged)
```

```{r, fig.align='center', fig.height=5, fig.width=7}
DimPlot(merged,  group.by = c("batch"), shuffle = T)
```

Clear batch effect. 

## Integration: Harmony

```{r}
merged <- merge(yap1_wildtype,
                y = meox1_wildtype,
                project = "wildtype_only", merge.data = TRUE) #don't join layers, leave then in there

#run basic workflow to get PCA
merged <- NormalizeData(merged) %>% 
  FindVariableFeatures() %>% 
  ScaleData() %>% 
  RunPCA()

```

```{r}
merged <- IntegrateLayers(object = merged, method = HarmonyIntegration, orig.reduction = "pca",
  new.reduction = 'harmony', verbose = FALSE)

```

Run workflow:

```{r}
merged <- RunUMAP(merged, dims = 1:30, reduction = "harmony") %>% #defaults
    FindNeighbors(., dims = 1:30, reduction = "harmony") %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1), reduction = "harmony")
```

## PCA correlations

```{r}
run_harmony_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$harmony@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, harmony_1, harmony_2, harmony_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "harmony_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "harmony_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_2") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "harmony_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_3") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```

```{r, fig.align='center', fig.height=5, fig.width=15}
wrap_plots(run_harmony_correlations(merged), ncol = 3)
```


## UMAP overview

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("batch"), shuffle = T) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```


```{r, fig.align='center', fig.height=10, fig.width=12}
wrap_plots(plot_umap_dataset(merged), ncol = 2)
```

#### Barplot for batch per cluster

```{r, fig.align='center', fig.height=5, fig.width=10}
merged@meta.data %>% 
  group_by(batch, seurat_clusters) %>% 
  summarise(., count = n()) %>% 
  ggplot(., aes(y = count, x= seurat_clusters, fill= batch))+
  geom_bar(stat = "identity", position = 'fill')+
  plot_sizing
```

## Low quality clusters

```{r, fig.align='center', fig.height=5, fig.width=15}
plot_list <- list()
plot_list[[1]] <- merged@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
plot_list[[2]] <- merged@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")


wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

## Clustree

## scGate: Find endothelial subset

```{r}
endothelial_gating_model <- gating_model(name = "endothelial", signature = c("kdrl+", 
                                                                    "col1a1a-",
                                                                    "epcam-",
                                                                    "gata2b-", 
                                                                    "myod1-", 
                                                                    "gch2-",
                                                                    "aox5-"))

merged <- scGate(data = merged,
                   model = endothelial_gating_model,
                   output.col.name = "is_endothlial")

```

```{r, fig.align='center', fig.height=5, fig.width=7}
pure_colours <- c("grey", "green3"); names(pure_colours) <- c("Impure", "Pure")
DimPlot(merged, group.by = "is_endothlial", cols = pure_colours)
```

Make purity plots for res 0.2:

```{r, fig.align='center', fig.height=5, fig.width=8}
merged@meta.data %>% 
  group_by(., RNA_snn_res.0.2, is_endothlial) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.2) %>%
    mutate(., RNA_snn_res.0.2_cluster_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.2_cluster_purity) %>% distinct() %>% 
    ggplot(., aes(x= reorder(RNA_snn_res.0.2, -RNA_snn_res.0.2_cluster_purity),  y = RNA_snn_res.0.2_cluster_purity))+
    geom_bar(stat = "identity") + plot_sizing +
    ylab("Endo Purity per cluster") + xlab("cluster") 
```

## DotPlot

Use some known marker genes for celltypes to double check scGate annotation.

```{r}
dotplot.level.01 <- c(
  "cdh6",
  "prox1a",
  "stab2",
  "lyve1b",
  "mrc1a",
  "flt1",
  "dll4",
  "esm1",
  "hlx1",
  "hand2",
  "fn1a",
  "gata2b",
  "runx1",
  "alas2",
  "blvrb",
  "spi1b",
  "mpx",
  "neurod4",
  "nova2",
  "cdh1",
  "epcam",
  "myod1",
  "myog",
  "pmela",
  "dct",
  "lum",
  "col1a1a",
  "krt4",
  "cyt1",
  "cryba1b",
  "crybb1",
  "ednrba",
  "pnp4a",
  "gch2",
  "aox5")
```

Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(group)  +plot_sizing_seurat + rotate_x
}

```

Make dotplots for all clusters

```{r, fig.align='center', fig.height=5, fig.width=7}

DimPlot(merged, group.by = "RNA_snn_res.0.2", label = T)
```

```{r, fig.align='center', fig.height=5, fig.width=10}
run_dotplot(merged, dotplot.level.01, group = "RNA_snn_res.0.2")

```
Make dotplot for clusters that have mid-range purity.

```{r}
run_dotplot_mid_purity <- function(seurat, genes, limits){
  clusters <- seurat@meta.data %>% 
    group_by(., RNA_snn_res.0.2, is_endothlial) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.2) %>%
    mutate(., RNA_snn_res.0.2_cluster_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.2_cluster_purity) %>% distinct() %>% 
    filter(., RNA_snn_res.0.2_cluster_purity > limits[1] & RNA_snn_res.0.2_cluster_purity < limits[2]) %>% 
    pull(RNA_snn_res.0.2) %>% unique()
  
  seurat <- SetIdent(seurat, value = "RNA_snn_res.0.2")
  DotPlot(seurat, features = genes, group.by = "RNA_snn_res.0.2", idents = clusters)  +plot_sizing_seurat + rotate_x
}

```

```{r, fig.align='center', fig.height=4, fig.width=10}
run_dotplot_mid_purity(merged, limits = c(0.2, 0.8), genes = dotplot.level.01)
```

They all express other markers as well. 


## Export

```{r}

file_name <- here('output/data/Datsets/merged_wildtypes/merged_wildtypes_Level_01_annotated.qs2')
qs_save(merged, file_name)

```




