---
title: "Combined WT only dataset: Make Level 03 (LEC/VEC) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 03 object for yap1 dataset. 

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_combining_wildtypes_dataset_makeObject_Level02_annotation.Rmd

```{r}
level_02 <- qs_read(here('output/data/Datsets/merged_wildtypes/merged_wildtypes_Level_02_annotated.qs2'))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: LEC/VEC clusters only

Subset Level 01 to only include LEC and VEC subsets. Annotation was done in .pre_processing_combining_wildtypes_dataset_makeObject_Level02_annotation.Rmd

-> note that we are not keeping mixed EC as they express AEC markers, but we have a cycling LEC/VEC population we kept.

```{r}
celltype_to_keep <- c("LEC", "VEC", "cycling_LEC_VEC", "pre_muLEC", "hippo+_VEC")

run_subset_lecvec <- function(seurat, keep = celltype_to_keep){
  cells_to_keep <- seurat@meta.data %>% 
  filter(., L2_celltype %in% celltype_to_keep) %>% 
  rownames_to_column("Barcode") %>% 
  pull(Barcode) 
  
  seurat <- subset(seurat, cells = cells_to_keep)
  return(seurat)
}
```


```{r}
level_03 <- run_subset_lecvec(level_02)
  
```

Clean up list, remove Level 02

```{r}
rm(level_02); gc()
```


## Re-run single-cell workflow


Not using sctransform, regressing out %mito. Same function we used for Level 01, 02

```{r, message=FALSE, warning=FALSE}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #this is technically unneccessary, but also won't hurt
    FindVariableFeatures() %>% 
    ScaleData(., split.by = "batch") %>% 
    RunPCA() %>% 
    IntegrateLayers(object = ., method = HarmonyIntegration, orig.reduction = "pca",
  new.reduction = 'harmony', verbose = FALSE) %>% 
    RunUMAP(., dims = 1:30, reduction = "harmony") %>% #defaults
    FindNeighbors(., dims = 1:30, reduction = "harmony") %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1), reduction = "harmony")
}
```


```{r}
level_03 <- run_workflow(level_03)
```


## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=5, fig.width=15}
wrap_plots(run_pca_correlations(level_03), ncol = 3)
```

Correlation lib size with harmony 1

## UMAP overviews

Make UMAPs for features, ncounts, clusters, genotype, cell cycle, Level 2 celltypes

```{r}
plot_umap_dataset <- function(seurat){
  colours <- Polychrome::dark.colors() %>% unname()
  
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters"), label = T) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("batch"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[5]] <- DimPlot(seurat,  group.by = c("tricycle_stage"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[6]] <- DimPlot(seurat,  group.by = c("L2_celltype"), shuffle = T,cols =colours) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=15, fig.width=12}

wrap_plots(plot_umap_dataset(level_03), ncol=2)
```

## Clustree + Composition


```{r, fig.align='center', fig.height=6, fig.width=6}
clustree(level_03)
```

Stacked barplot for different resolutions capturing genotype split.

```{r, fig.align='center', fig.height=8, fig.width=25}
level_03@meta.data %>% 
    select(., batch, starts_with("RNA")) %>% 
    pivot_longer(., cols = starts_with("RNA"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("RNA_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, batch) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= batch)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =2) + 
    strip_fix 
```


## Low quality cluster - meox1 dataset

We know that the meox1 dataset has these weird low count cells that were not picked up at higher levels because they mixed nicely with other clusters. Is that one of the meox1 populations that stands out?


```{r, fig.align='center', fig.height=5, fig.width=15}
plot_list <- list()
plot_list[[1]] <- level_03@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")

plot_list[[2]] <- level_03@meta.data %>% 
  ggplot(., aes(y = log10(nFeature_RNA), x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
plot_list[[3]] <- level_03@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")


wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```


One cluster does stand out a bit, cluster 4 in seurat clusters. Is that a meox1 enriched cluster?

```{r, fig.align='center', fig.align='center', fig.height=5, fig.width=8}
level_03@meta.data %>% 
    select(., batch, seurat_clusters) %>% 
    group_by(seurat_clusters,batch) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= seurat_clusters, y = count, fill= batch)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing + ggtitle("seurat clusters composition")
```

Yep, cluster 4 is a bit of a meox1 cluster...

## Feature UMAPs


```{r}
level_03_dotplot <- c(
  "dll4", #aec
  "flt1", #aec
  "stab2",  #cvp
  "stab1", #cvp
  "ifi30",#cvp
  "mafba",#cvp
  "cxcl12a", #mapk signaling
  "esm1", #migrating
  "hlx1", #migrating
  "mki67", #cycling
  "pcna", #cycling
  "nop53", #lec? 
  "ddx21", #lec?
  "tbx1", #lec
  "prox1a", #LEC
  "cdh6", #LEC, hippo
  "fn1a", #hippo
  "osr2", #muLEC
  "sparc"
)
```

```{r}
run_feature_umap <- function(seurat, genes){
  FeaturePlot(seurat, features = genes, ncol = 4) 
}

```

```{r, fig.align='center', fig.height=20, fig.width=20}
run_feature_umap(seurat = level_03, genes = level_03_dotplot)
```

Small selection:

```{r, fig.align='center', fig.height=15, fig.width=10}
FeaturePlot(level_03, features = c("prox1a", "osr2", "ifi30", "fn1a", "cdh6"), ncol = 2) 

```

So interestingly, the meox1 cluster with low counts seems to be a hippo+ cluster...

Small selection #2:

```{r, fig.align='center', fig.height=10, fig.width=5}
FeaturePlot(level_03, features = c("dll4", "lyve1b"), ncol = 1) 

```

## Compare GEX to time course data

```{r}
Fish_atlas <- readRDS("/Users/meiermichelle/Library/CloudStorage/OneDrive-PeterMac/HOGAN_LAB/analysis/GeneralDataset/Level_02_D10042_LymphaticTimeCourse_annotated.RDS")
Fish_atlas$Stage <- factor(Fish_atlas$Stage, levels = c("40hpf", "3dpf", "4dpf", "5dpf", "7dpf"))

```

Make dotplots

```{r}
run_dotplot_atlas <- function(genes, max_expression){
  plot_list <- lapply(levels(Fish_atlas$Stage), function(stage){
  subset_stage <- subset(Fish_atlas, cells = colnames(Fish_atlas)[Fish_atlas$Stage == stage])
  Idents(subset_stage) <- "L2_Predicted_phenotype"
  DotPlot(subset_stage, features = genes, group.by = "L2_Predicted_phenotype", scale.min = 0,
          scale.max = max_expression, idents = c("LEC", "VEC"), scale = F)+ RotatedAxis() + ylab("") + xlab("") + ggtitle(stage)})
   wrap_plots(plot_list, nrow = 1) + plot_layout(guides = "collect", axes = "collect")

  }

```

```{r, fig.align='center', fig.height=5, fig.width=10}
run_dotplot_atlas("ifi30", max_expression = 60)
```




## Dotplot for genes


Make dotplot function:


```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(sprintf("%s, %s",group, seurat@misc$name))  +plot_sizing_seurat + rotate_x
}

```


```{r, fig.align='center', fig.height=5, fig.width=8}
run_dotplot(seurat = level_03, genes = level_03_dotplot, group = "seurat_clusters")
```



## pre-muLECs?

Look at % expressed for osr2+ cluster

```{r}
plot_info <- DotPlot(level_03, features = "osr2", split.by = "batch", group.by = "seurat_clusters")

plot_info@data %>% filter(., grepl("12_", id)) %>% select(, avg.exp, pct.exp, id)
```



## Add annotations

Can't commit yet.

## Export 

Export Level 03. 

```{r}

file_name <- here('output/data/Datsets/merged_wildtypes/merged_wildtypes_Level_03.qs2')
qs_save(level_03, file_name)

```

