---
title: "D10025 yap1 dataset: Make Level 01 (all cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 01 object for yap1 dataset. 

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
})
```

### Load general info

List of ribosomal and globin genes that we're going to remove

```{r}
ribo_genes <- read.table(here("data/genesets/DR_ribosomal_globin_genes.csv"), sep="\t", stringsAsFactors=FALSE)

```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```

## Load data

Read in 10X outputs.

S20127 = yap1 mutant

S20128 = wildtype

```{r}
yap1_object <- Read10X(here("data/input/S20127/filtered_feature_bc_matrix"))
wt_object <- Read10X(here("data/input/S20128/filtered_feature_bc_matrix"))
```

Remove ribo and globin genes

```{r}
yap1_object <- yap1_object[!rownames(yap1_object) %in% ribo_genes$gene_name,]
wt_object <- wt_object[!rownames(wt_object) %in% ribo_genes$gene_name,]
```

## Make Seurat object

```{r}
yap1_seurat <- CreateSeuratObject(counts=yap1_object,
													project = "S20127",
													min.cells = 5)

#add genotype
yap1_seurat$Genotype <- "yap1_mutant"

wt_seurat <- CreateSeuratObject(counts=wt_object,
													project = "S20128",
													min.cells = 5)

wt_seurat$Genotype <- "wildtype"

```

Combine them into list to use lapply later. 

```{r}
list_objects <- list(yap_mutant = yap1_seurat, 
                     wt = wt_seurat)
```

Clean up

```{r, message=FALSE, warning=FALSE}
rm(yap1_object, yap1_seurat, wt_object, wt_seurat, ribo_genes); gc()
```

## % mitochondria

```{r}
list_objects <- lapply(list_objects, function(seurat){
  seurat$`percent.mt` <- PercentageFeatureSet(seurat,
												pattern = "^[mtMtMT-]")
  return(seurat)
})

```


## Pre-QC metrics

Make function for density plots and correlation plots

```{r}
make_density_plots <- function(seurat){
  plot_list <- list()

  #ncount
  plot_list[[1]] <- ggplot(seurat@meta.data, aes(x=log10(nCount_RNA))) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('log10(count)') +
    plot_sizing + ggtitle(sprintf("Library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggplot(seurat@meta.data, aes(x=log10(nFeature_RNA))) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('log10(nfeature)') +
    plot_sizing + ggtitle(sprintf("#features, %s",  seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggplot(seurat@meta.data, aes(x=percent.mt)) + 
    geom_histogram(aes(y=..density..), colour="black", fill="white")+
    geom_density(alpha=.2, fill="#FF6666")+ xlab('percent.mt') +
    plot_sizing + ggtitle(sprintf("mito, %s",  seurat@project.name))
  
  return(plot_list)
  
}

make_correlation_plots <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "log10ncount", y = "log10nfeat",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(#features)") + xlab("log10(nCount)") +
    stat_cor(method = "pearson", label.x = 2.5, label.y = 4, size = 5) +
    ggtitle(sprintf("Lib size vs #features, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "log10ncount", y = "percent.mt",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("%mito") + xlab("log10(nCount)") +
    stat_cor(method = "pearson", label.x = 3.5, label.y = 80, size = 5) +
    ggtitle(sprintf("Lib size vs mito, %s",  seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "log10nfeat", y = "percent.mt",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("%mito") + xlab("log10(#features)") +
    stat_cor(method = "pearson", label.x = 3.5, label.y = 80, size = 5) +
    ggtitle(sprintf("#features vs mito, %s",  seurat@project.name))
  
  return(plot_list)
  
}
```


### Library size, nFeatures, %mito

Density plots

```{r, fig.align='center', fig.height=8, fig.width=15}
plot_list <- lapply(list_objects, make_density_plots)

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect")
```

Small hump, still reasonably large library size there though.


### Correlations


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, make_correlation_plots)

wrap_plots(unlist(plot_list, recursive = F)) 
```

## Doublet calling

Using scDblFinder. No QC before running this as recommended. 

```{r}
run_scDblFinder <- function(seurat){
  sce <- as.SingleCellExperiment(seurat) #convert
  
  sce <- logNormCounts(sce) %>% #run steps
  runPCA() %>%
  runUMAP()
  
  set.seed(1997)#for reproducibility
  sce <- scDblFinder(sce, clusters = T) #run scDblFinder with clusters
  
  #add to seurat
  seurat$scDblFinder.class <- sce$scDblFinder.class #actual label
  seurat$scDblFinder.cxds_score <- sce$scDblFinder.cxds_score #co-expression score
  
  return(seurat)
}
```

Run

```{r}
list_objects <- lapply(list_objects, run_scDblFinder)
```

Around 20% doublets, predicted. 10X doublet rate is 0.008 per 10k. 

```{r}
#expected doublet rate
(ncol(list_objects$yap_mutant)* 0.008)/1000
(ncol(list_objects$wt)* 0.008)/1000
```


```{r}
#observed
table(list_objects$yap_mutant$scDblFinder.class)/ncol(list_objects$yap_mutant)
table(list_objects$wt$scDblFinder.class)/ncol(list_objects$wt)
```

Slightly higher than expected for mutant. Importantly, this is a timepoint (2dpf) where cells are transitioning between celltypes and may be co-expressing genes which could through it off. 

Library size and nFeature difference between doublets and singlet:

```{r}
run_metrics_dbl <- function(seurat){
  plot_list <- list()
  
  plot_list[[1]] <- seurat@meta.data %>% 
    ggplot(., aes(y = log10(nCount_RNA), x = scDblFinder.class)) +
    geom_boxplot()+ plot_sizing + ylab("log10(count)") + xlab("") +
    ggtitle(seurat$Genotype)
  
  plot_list[[2]] <- seurat@meta.data %>% 
    ggplot(., aes(y = log10(nFeature_RNA), x = scDblFinder.class)) +
    geom_boxplot()+ plot_sizing + ylab("log10(features)") + xlab("") +
    ggtitle(seurat$Genotype)
  
  return(plot_list)
}

```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_metrics_dbl)

wrap_plots(unlist(plot_list, recursive = F)) 
```

## Explore and remove outliers

%mito: Fixed threshold 20%

Library size: nmads 2, lower and higher

nFeatures: nmads 2, lower and higher

### Library size

Get library size threshold for each sample - lower bound

```{r}
#yap mutant
lib_size_outliers_low <- isOutlier(list_objects$yap_mutant$nCount_RNA, type = "lower", nmads = 2, log = T)  
lib_size_outliers_low %>% sum()
summary(list_objects$yap_mutant$nCount_RNA[lib_size_outliers_low]) 

```

```{r}
#wildtype
lib_size_outliers_low <- isOutlier(list_objects$wt$nCount_RNA, type = "lower", nmads = 2, log = T)  
lib_size_outliers_low %>% sum()
summary(list_objects$wt$nCount_RNA[lib_size_outliers_low]) 

```

Very similar

Upper bound ncounts

```{r}
#yap mutant
lib_size_outliers_upper <- isOutlier(list_objects$yap_mutant$nCount_RNA, type = "higher", nmads = 2, log = T)  
lib_size_outliers_upper %>% sum()
summary(list_objects$yap_mutant$nCount_RNA[lib_size_outliers_upper]) 

list_objects$yap_mutant$scDblFinder.class[lib_size_outliers_upper] %>% table()
```

```{r}
#wildtype
lib_size_outliers_upper <- isOutlier(list_objects$wt$nCount_RNA, type = "higher", nmads = 2, log = T)  
lib_size_outliers_upper %>% sum()
summary(list_objects$wt$nCount_RNA[lib_size_outliers_upper]) 

list_objects$wt$scDblFinder.class[lib_size_outliers_upper] %>% table()
```

Combined

```{r}
#yap mutant
lib_size_outliers_combined <- isOutlier(list_objects$yap_mutant$nCount_RNA, type = "both", nmads = 2, log = T)  
lib_size_outliers_combined %>% sum()
```

```{r}
#wildtype
lib_size_outliers_combined <- isOutlier(list_objects$wt$nCount_RNA, type = "both", nmads = 2, log = T)  
lib_size_outliers_combined %>% sum()
```

### Features

```{r}
#yap mutant
lib_size_outliers_low <- isOutlier(list_objects$yap_mutant$nFeature_RNA, type = "lower", nmads = 2, log = T)  
lib_size_outliers_low %>% sum()
summary(list_objects$yap_mutant$nFeature_RNA[lib_size_outliers_low]) 

```

```{r}
#wildtype
lib_size_outliers_low <- isOutlier(list_objects$wt$nFeature_RNA, type = "lower", nmads = 2, log = T)  
lib_size_outliers_low %>% sum()
summary(list_objects$wt$nFeature_RNA[lib_size_outliers_low]) 

```


Upper bound ncounts

```{r}
#yap mutant
lib_size_outliers_upper <- isOutlier(list_objects$yap_mutant$nFeature_RNA, type = "higher", nmads = 2, log = T)  
lib_size_outliers_upper %>% sum()
summary(list_objects$yap_mutant$nFeature_RNA[lib_size_outliers_upper]) 

list_objects$yap_mutant$scDblFinder.class[lib_size_outliers_upper] %>% table()
```

```{r}
#wildtype
lib_size_outliers_upper <- isOutlier(list_objects$wt$nFeature_RNA, type = "higher", nmads = 2, log = T)  
lib_size_outliers_upper %>% sum()
summary(list_objects$wt$nFeature_RNA[lib_size_outliers_upper]) 

list_objects$wt$scDblFinder.class[lib_size_outliers_upper] %>% table()
```

Combined

```{r}
#yap mutant
lib_size_outliers_combined <- isOutlier(list_objects$yap_mutant$nFeature_RNA, type = "both", nmads = 2, log = T)  
lib_size_outliers_combined %>% sum()
```

```{r}
#wildtype
lib_size_outliers_combined <- isOutlier(list_objects$wt$nFeature_RNA, type = "both", nmads = 2, log = T)  
lib_size_outliers_combined %>% sum()
```

### Remove outliers

```{r}
remove_outliers <- function(seurat){
  mito <- seurat$percent.mt > 20
  features <-isOutlier(seurat$nFeature_RNA, type = "both", nmads = 2, log = T)
  count <-isOutlier(seurat$nCount_RNA, type = "both", nmads = 2, log = T)
  
  remove_vec <- c(mito | features | count)
  
  seurat <- seurat[, !remove_vec]
}
```

```{r}
list_objects <- lapply(list_objects, remove_outliers)
```

How many doublets are left?

```{r}
list_objects$yap_mutant$scDblFinder.class %>% table()
list_objects$wt$scDblFinder.class %>% table()
```


## Option: remove doublets

I think we should at least try to see if removing doublets changes outcome, though I still think there is a real chance intermediate stages of endothelial cells will be marked as doublets. 

Add removed doublets object to list. 

```{r}
list_objects$yap_mutant_nodoublets <- list_objects$yap_mutant[,list_objects$yap_mutant$scDblFinder.class == "singlet"]

list_objects$wt_nodoublets <- list_objects$wt[,list_objects$wt$scDblFinder.class == "singlet"]

```

```{r}
list_objects %>% length()
```


## Post-QC metrics


### Library size, nFeatures, %mito

Density plots

With doublets

```{r, fig.align='center', fig.height=8, fig.width=15}
plot_list <- lapply(list_objects[1:2], make_density_plots)

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") +plot_annotation(title = "with doublets")
```

Without doublets

```{r, fig.align='center', fig.height=8, fig.width=15}
plot_list <- lapply(list_objects[3:4], make_density_plots)

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") +plot_annotation(title = "without doublets")
```

General distribution without doublets looks slightly better.

### Correlations


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects[1:2], make_correlation_plots)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "with doublets")
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects[3:4], make_correlation_plots)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "without doublets")
```

No real differences here.

## Run single-cell workflow

```{r, include=FALSE}
rm(plot_list);gc()
```


We're going to run the workflow on each sample separately (just in case we need it) and on combined objects. 

Not using sctransform, regressing out %mito

```{r}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #should deal with basic library size differences
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1))
}
```


Combine the samples

```{r}
list_object_combined <- list()
list_object_combined$merged <- merge(list_objects$yap_mutant,
                                     y = list_objects$wt,
                                     project = "Dataset_10025", merge.data = TRUE) 
list_object_combined$merged <- JoinLayers(list_object_combined$merged) #seurat v5 things

list_object_combined$merged_nodoublets <- merge(list_objects$yap_mutant_nodoublets,
                                     y = list_objects$wt_nodoublets,
                                     project = "Dataset_10025", merge.data = TRUE)
list_object_combined$merged_nodoublets <- JoinLayers(list_object_combined$merged_nodoublets) #seurat v5 things

  
```

### Each Sample


```{r, message=FALSE}
list_objects <- lapply(list_objects, run_workflow)
```


### Combined


```{r, message=FALSE}
list_object_combined <- lapply(list_object_combined, run_workflow)
```

## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects[1:2], run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "with doublets")
```

Combined (top with, bottom without doublets):

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_object_combined, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "with doublets")

```

No bias.

## UMAP overviews

### Each Sample

Not that much to see, only make UMAPs for nfeature, ncount and seurat_cluster.

```{r}
plot_umap_sample <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects[1:2], plot_umap_sample)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "with doublets")
```

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects[3:4], plot_umap_sample)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "without doublets")
```

### Combined

Not that much to see, only make UMAPs for nfeature, ncount and seurat_cluster.

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=20}
plot_list <- lapply(list_object_combined, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F)) +plot_annotation(title = "without doublets") + plot_layout(nrow = 2)
```

### Endothelial cell genes

Quick overview of endothelial cell genes.

```{r}
run_umap_genes <- function(seurat, genes = c("cdh5", "kdrl", "cdh6", "prox1a")){
  p <- FeaturePlot(seurat, features = genes) 
  p + plot_annotation(title = seurat@project.name)
}
```

#### Each Sample

With doublets

```{r, fig.align='center', fig.height=10, fig.width=15}
lapply(list_objects[1:2], run_umap_genes)
```

Without doublets

```{r, fig.align='center', fig.height=10, fig.width=15}
lapply(list_objects[3:4], run_umap_genes)
```

#### Combined

With doublets:

```{r, fig.align='center', fig.height=10, fig.width=15}
run_umap_genes(list_object_combined$merged)
```

Without doublets:

```{r, fig.align='center', fig.height=10, fig.width=15}
run_umap_genes(list_object_combined$merged_nodoublets)
```


## Export 

Export them all.

```{r}
#dataset
file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01.qs2')
qs_save(list_object_combined$merged_nodoublets, file_name)
file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_kept_doublets.qs2')
qs_save(list_object_combined$merged, file_name)

#samples
file_name <- here('output/data/Samples/S20127/S20127_yap1_mutant_Level_01.qs2')
qs_save(list_objects$yap_mutant_nodoublets, file_name)
file_name <- here('output/data/Samples/S20127/S20127_yap1_mutant_Level_01_kept_doublets.qs2')
qs_save(list_objects$yap_mutant, file_name)

file_name <- here('output/data/Samples/S20128/S20128_wildtype_Level_01.qs2')
qs_save(list_objects$wt_nodoublets, file_name)
file_name <- here('output/data/Samples/S20128/S20128_wildtype_Level_01_kept_doublets.qs2')
qs_save(list_objects$wt, file_name)


```


