---
title: "D10025 yap1 dataset Level 03 vs D10051 meox1 dataset Level 03 - correlation"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Check overlap of transcriptomic profiles of clusters between the two datasets.


### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
  library(bluster)
})
```


### Load data

Made here: ./pre_processing_D10025_yap1_dataset_makeObject_Level03.Rmd and ./pre_processing_D10051_meox1_dataset_makeObject_Level03_stringent.Rmd

Only using the ones where doublets were removed.

```{r}
input_paths <- c(file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_03.qs2'),
                 here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_03_stringent.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_keeplegend <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16)) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Correlation

### Re-scale with all genes

```{r}
run_scale_all_genes <- function(seurat){
  seurat <- ScaleData(seurat, features = rownames(seurat))
  return(seurat)
}

list_objects <- lapply(list_objects, run_scale_all_genes)
```

### Run correlation - clusters

Run correlation between the two datasets.

```{r}
run_cluster_correlation <- function(seurat_yap = list_objects$D10025_yap1_dataset_Level_03,
                                    seurat_meox = list_objects$D10051_meox1_dataset_Level_03_stringent,
                                    resolution_yap = "RNA_snn_res.0.1",
                                    resolution_meox = "RNA_snn_res.0.1"){
  #get GEX
  Expression_assay_yap <- GetAssayData(seurat_yap, layer = "scale.data", assay = "RNA")
  Expression_assay_meox <- GetAssayData(seurat_meox, layer = "scale.data", assay = "RNA")
  #intersect of genes
  genes <- intersect(rownames(Expression_assay_yap), rownames(Expression_assay_meox))
  Expression_assay_yap <- Expression_assay_yap[genes,]
  Expression_assay_meox <- Expression_assay_meox[genes,]
  
  #get cluster info
  clusters_yap <- seurat_yap@meta.data %>% 
      pull(resolution_yap) %>% unique()
  clusters_meox <- seurat_meox@meta.data %>% 
      pull(resolution_meox) %>% unique()
  
  #initiate matrices
  corr_mat <- pval_mat <- matrix(data = NA, nrow = length(clusters_yap), ncol = length(clusters_meox))
  colnames(corr_mat) <- colnames(pval_mat) <- clusters_meox
  rownames(corr_mat) <- rownames(pval_mat) <- clusters_yap
    
    #Populate correlation matrix
  for (cluster1 in clusters_yap){
    for (cluster2 in clusters_meox){
      #get barcodes
      barcodes1 <- seurat_yap@meta.data %>% 
        rownames_to_column("Barcode") %>% 
        dplyr::rename(chosen_res=resolution_yap) %>% 
        filter(., chosen_res == cluster1) %>% pull(Barcode)
      barcodes2 <- seurat_meox@meta.data %>% 
        rownames_to_column("Barcode") %>% 
        dplyr::rename(chosen_res=resolution_meox) %>% 
        filter(., chosen_res == cluster2) %>% pull(Barcode)
        
      #filter data sets
      expr_yap <- Expression_assay_yap[, colnames(Expression_assay_yap) %in% barcodes1] %>% as.data.frame() %>% 
        apply(., 1, mean)
      expr_meox <- Expression_assay_meox[, colnames(Expression_assay_meox) %in% barcodes2] %>% as.data.frame() %>% 
        apply(., 1, mean)

      #calculate correlation
        cor_res <- cor.test(expr_yap, expr_meox, method = 'spearman')
        corr_mat[cluster1, cluster2] <- cor_res$estimate
        pval_mat[cluster1, cluster2] <- cor_res$p.value
        
      } 
    
  }

  return(list(corr=corr_mat, pval=pval_mat))

}
```


```{r, message=FALSE}
result <- run_cluster_correlation(resolution_yap = "RNA_snn_res.0.2", 
                                  resolution_meox = "RNA_snn_res.0.2")
```



```{r}
run_correlation_plots <- function(list_in){
    #get correlation ready
    plot_df_corr <- list_in$corr %>% as.data.frame() %>%
      rownames_to_column("cluster1") %>% 
    pivot_longer(!cluster1, names_to = "cluster2", values_to = "Correlation") %>% 
      mutate(., cluster1= factor(cluster1, levels = as.character(seq(0,100,1))), #overshoot
                cluster2= factor(cluster2, levels = as.character(seq(0,100,1))))
    #get pvals ready
    plot_df_pval <- list_in$pval %>% as.data.frame() %>%
      rownames_to_column("cluster1") %>% 
    pivot_longer(!cluster1, names_to = "cluster2", values_to = "p-value") %>% 
      mutate(., cluster1= factor(cluster1, levels = as.character(seq(0,100,1))), #overshoot
                cluster2= factor(cluster2, levels = as.character(seq(0,100,1))))
    plot_df <- full_join(plot_df_corr,plot_df_pval) %>% 
      mutate(., signs = ifelse(`p-value` < 0.05, round(Correlation, 2), "x"))
    
    #make plot
    heatmap.correlation <- ggplot(plot_df, 
                                  aes(x = cluster1, y = cluster2, fill = Correlation)) +
      geom_tile() + theme_bw() +
      scale_fill_gradient2(high = '#762A83', low = "#1B7837", mid = 'white', limits =c(-1,1)) +
      geom_text(aes(label = signs)) +
      labs(fill = 'Correlation Coefficient') +
      plot_sizing_keeplegend +
      xlab("yap1 clusters") + ylab("meox1 clusters")
    
    return(heatmap.correlation)
  
}
```


```{r, fig.align='center', fig.height=10, fig.width=10}
run_correlation_plots(result)
```

### Run correlation - temporary celltype labels

Add temporary celltype labels to make it easier to read.

```{r}
#yap1
list_objects$D10025_yap1_dataset_Level_03$L3_celltype <- case_when(
  list_objects$D10025_yap1_dataset_Level_03$RNA_snn_res.0.2 == 5 ~ "yap1_mutant",
  list_objects$D10025_yap1_dataset_Level_03$RNA_snn_res.0.2 == 4 ~ "LEC",
  list_objects$D10025_yap1_dataset_Level_03$RNA_snn_res.0.2 == 3 ~ "cvpVEC",
  list_objects$D10025_yap1_dataset_Level_03$RNA_snn_res.0.2 == 2 ~ "hippoVEC",
  list_objects$D10025_yap1_dataset_Level_03$RNA_snn_res.0.2 == 1 ~ "msVEC",
  list_objects$D10025_yap1_dataset_Level_03$RNA_snn_res.0.2 == 0 ~ "preLEC"
)

#meox1
list_objects$D10051_meox1_dataset_Level_03_stringent$L3_celltype <- case_when(
  list_objects$D10051_meox1_dataset_Level_03_stringent$RNA_snn_res.0.2 == 3 ~ "pre_muLEC",
  list_objects$D10051_meox1_dataset_Level_03_stringent$RNA_snn_res.0.2 == 2 ~ "hippoVEC",
  list_objects$D10051_meox1_dataset_Level_03_stringent$RNA_snn_res.0.2 == 1 ~ "LEC",
  list_objects$D10051_meox1_dataset_Level_03_stringent$RNA_snn_res.0.2 == 0 ~ "msVEC"
)

```

Re-run correlation

```{r, message=FALSE}
result <- run_cluster_correlation(resolution_yap = "L3_celltype", 
                                  resolution_meox = "L3_celltype")
```

Make new plotting function, add levels

```{r}
run_correlation_plots <- function(list_in, selected_levels){
    #get correlation ready
    plot_df_corr <- list_in$corr %>% as.data.frame() %>%
      rownames_to_column("cluster1") %>% 
    pivot_longer(!cluster1, names_to = "cluster2", values_to = "Correlation") %>% 
      mutate(., cluster1= factor(cluster1, levels = selected_levels), #overshoot
                cluster2= factor(cluster2, levels = selected_levels))
    #get pvals ready
    plot_df_pval <- list_in$pval %>% as.data.frame() %>%
      rownames_to_column("cluster1") %>% 
    pivot_longer(!cluster1, names_to = "cluster2", values_to = "p-value") %>% 
      mutate(., cluster1= factor(cluster1, levels = selected_levels), #overshoot
                cluster2= factor(cluster2, levels = selected_levels))
    plot_df <- full_join(plot_df_corr,plot_df_pval) %>% 
      mutate(., signs = ifelse(`p-value` < 0.05, round(Correlation, 2), "x"))
    
    #make plot
    heatmap.correlation <- ggplot(plot_df, 
                                  aes(x = cluster1, y = cluster2, fill = Correlation)) +
      geom_tile() + theme_bw() +
      scale_fill_gradient2(high = '#762A83', low = "#1B7837", mid = 'white', limits =c(-1,1)) +
      geom_text(aes(label = signs)) +
      labs(fill = 'Correlation Coefficient') +
      plot_sizing_keeplegend +
      xlab("yap1 clusters") + ylab("meox1 clusters")
    
    return(heatmap.correlation)
  
}
```


```{r, fig.align='center', fig.height=5, fig.width=8}
levels_to_use <- c("hippoVEC", "msVEC", "cvpVEC", "preLEC" , "LEC", "pre_muLEC", "yap1_mutant")
run_correlation_plots(result, selected_levels = levels_to_use) + rotate_x + ggtitle("yap1 vs meox1 clusters")
```

How do these correlation values compare to within-dataset correlation? Particularly for the yap1 dataset.

Highjack function

```{r, message=FALSE}
result_within_yap <- run_cluster_correlation(seurat_yap = list_objects$D10025_yap1_dataset_Level_03,
                                             seurat_meox =  list_objects$D10025_yap1_dataset_Level_03,
                                             resolution_yap = "L3_celltype", 
                                             resolution_meox = "L3_celltype")
```

Make new plotting function, add levels


```{r, fig.align='center', fig.height=5, fig.width=8}
levels_to_use <- c("hippoVEC", "msVEC", "cvpVEC", "preLEC" , "LEC", "yap1_mutant")
run_correlation_plots(result_within_yap, selected_levels = levels_to_use) + rotate_x + ylab("yap1 clusters") + ggtitle("yap1 datset: Within-dataset correlation")
```




