---
title: "Combined WT only dataset: Make Level 01 (all cells) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Combine wildtype from yap1 and meox1 experiments. 


### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
  library(tricycle)
 })
```

### Load data

From here ./pre_processing_D10025_yap1_dataset_makeObject_Level01.Rmd and here pre_processing_D10051_meox1_dataset_makeObject_Level01.Rmd. Note that each sample was processed there, e.g. QC was done on each sample separately.

Running with doublets removed: 

```{r}
yap1_wildtype <- qs_read(here("output/data/Samples/S20128/S20128_wildtype_Level_01.qs2"))
meox1_wildtype <- qs_read(here("output/data/Samples/S20201/S20201_wildtype_Level_01.qs2"))
```

Add batch information. 

```{r}
yap1_wildtype$batch <- "yap1_experiment"
meox1_wildtype$batch <- "meox1_experiment"
```

Read in mouse -> fish names

```{r}
mouse_to_fish <- qs_read(here("data/genesets/geneMap_mmusculus_to_drerio.qs2"))
```



### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Re-filter meox1 wildtype

See [here](https://atlassian.petermac.org.au/bitbucket/pages/HL/wflow_yap1_meox1_paper/master/browse/docs/pre_processing_D10051_meox1_dataset_makeObject_Level01_annotation.html), but the original filtering we performed did not really get rid of all the low quality cells, so we need to introduce a second filtering step where we remove cells with nFeatures < 400 (nCounts threshold already implemented is 600, both of these thresholds are informed through the yap1 dataset). 


```{r}
meox1_wildtype <- meox1_wildtype[, !meox1_wildtype$nFeature_RNA < 400]
```

## Downsample yap1 batch

yap1 batch is much larger, let's randomly downsample to match meox1 numbers

```{r}
yap1_wildtype <- yap1_wildtype[, sample(colnames(yap1_wildtype), size = ncol(meox1_wildtype), replace=F)]
```


## Tricycle: Run cell cycle analysis

Run tricycle on each separately as the merged layers later make it difficult.

Make function.

```{r}

run_tricycle_for_fish <- function(seurat, x_species_dictionary = mouse_to_fish){
  #prepare
  sce <- as.SingleCellExperiment(seurat, assay = "RNA") #convert to SCE
  sub_sce <- sce[rownames(sce) %in% x_species_dictionary$drerio_zfin_id_symbol,] #subset to genes in dic
  sub_x_species_dictionary <- x_species_dictionary %>%  #subset dictionary to only include the genes that we kept 
  filter(., drerio_zfin_id_symbol %in% rownames(sub_sce)) %>% 
  distinct(drerio_zfin_id_symbol, .keep_all = TRUE) #needed for function
  sub_x_species_dictionary <- sub_x_species_dictionary[match(rownames(sub_sce),
                                                             sub_x_species_dictionary$drerio_zfin_id_symbol),]
  stopifnot(all(sub_x_species_dictionary$drerio_zfin_id_symbol == rownames(sub_sce))) #check
  
  #run tricycle
  sub_sce <- project_cycle_space(sub_sce, 
                                 gname = sub_x_species_dictionary$mmusculus_mgi_symbol, 
                                 species = "mouse", 
                                 gname.type = "SYMBOL")
  #estimate position
  sub_sce <- estimate_cycle_position(sub_sce)
  
  
  #add back to seurat object
  seurat$tricycle_position <- sub_sce$tricyclePosition
  
  #add categorical cell cycle:
  # https://github.com/lb15/multiciliation_cycle/blob/main/mcc_timecourse/mcc_timecourse_analysis.R#L619
  seurat$tricycle_stage <- case_when(seurat$tricycle_position > pi/2 & seurat$tricycle_position < pi ~ "S",
                             seurat$tricycle_position >= pi & seurat$tricycle_position < pi*7/4 ~ "G2M",
                             TRUE ~ "G1/G0")
  
  rm(sce,sub_sce, sub_x_species_dictionary); gc()
  return(seurat)
  

}

```


```{r}
meox1_wildtype <- run_tricycle_for_fish(meox1_wildtype)
yap1_wildtype <- run_tricycle_for_fish(yap1_wildtype)

```

We know from processing the datasets independently that tricycle works, so won't run additional tests here.


## Combine 

Try to just combine

```{r}
merged <- merge(yap1_wildtype,
                y = meox1_wildtype,
                project = "wildtype_only", merge.data = TRUE) 

merged<- JoinLayers(merged) #seurat v5 things

```


Single-cell workflow:

Regress out %mito and batch

```{r}
run_workflow_batch <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #should deal with basic library size differences
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt", "batch")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = 0.8)
  return(seurat)
}

```

```{r}
merged <- run_workflow_batch(merged)
```

```{r, fig.align='center', fig.height=5, fig.width=7}
DimPlot(merged,  group.by = c("batch"), shuffle = T)
```

Clear batch effect. 

## Integration: Harmony

```{r, message=FALSE}
merged <- merge(yap1_wildtype,
                y = meox1_wildtype,
                project = "wildtype_only", merge.data = TRUE) #don't join layers, leave then in there

#run basic workflow to get PCA
merged <- NormalizeData(merged) %>% 
  FindVariableFeatures() %>% 
  ScaleData(., split.by = "batch") %>% #scale for each batch separately
  RunPCA()

```

```{r, message=FALSE}
merged <- IntegrateLayers(object = merged, method = HarmonyIntegration, orig.reduction = "pca",
  new.reduction = 'harmony', verbose = FALSE)

```

Run workflow:

```{r, message=FALSE}
merged <- RunUMAP(merged, dims = 1:30, reduction = "harmony") %>% #defaults
    FindNeighbors(., dims = 1:30, reduction = "harmony") %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1), reduction = "harmony")
```

## PCA correlations

```{r}
run_harmony_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$harmony@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, harmony_1, harmony_2, harmony_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "harmony_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "harmony_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_2") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "harmony_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("harmony_3") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("Harmony3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```

```{r, fig.align='center', fig.height=5, fig.width=15}
wrap_plots(run_harmony_correlations(merged), ncol = 3)
```


## UMAP overview

```{r}
plot_umap_dataset <- function(seurat){
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("batch"), shuffle = T) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```


```{r, fig.align='center', fig.height=10, fig.width=12}
wrap_plots(plot_umap_dataset(merged), ncol = 2)
```


## Low quality clusters

```{r, fig.align='center', fig.height=5, fig.width=15}
plot_list <- list()
plot_list[[1]] <- merged@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
plot_list[[2]] <- merged@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")


wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

## Clustree

```{r, fig.align='center', fig.height=10, fig.width=10}
clustree(merged)
```

Stacked barplot for different resolutions capturing genotype split.

```{r, fig.align='center', fig.height=15, fig.width=30}
merged@meta.data %>% 
    select(., batch, starts_with("RNA")) %>% 
    pivot_longer(., cols = starts_with("RNA"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("RNA_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, batch) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= batch)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =3) + 
    strip_fix 
```
## scGate: Find endothelial subset

```{r}
endothelial_gating_model <- gating_model(name = "endothelial", signature = c("kdrl+", 
                                                                    "col1a1a-",
                                                                    "epcam-",
                                                                    "gata2b-", 
                                                                    "myod1-", 
                                                                    "gch2-",
                                                                    "aox5-"))

merged <- scGate(data = merged,
                   model = endothelial_gating_model,
                   output.col.name = "is_endothlial")

```

```{r, fig.align='center', fig.height=5, fig.width=7}
pure_colours <- c("grey", "green3"); names(pure_colours) <- c("Impure", "Pure")
DimPlot(merged, group.by = "is_endothlial", cols = pure_colours)
```

Make purity plots for res 0.4:

```{r, fig.align='center', fig.height=5, fig.width=8}
merged@meta.data %>% 
  mutate(.,RNA_snn_res.0.4 = as.factor(RNA_snn_res.0.4)) %>% 
  group_by(., RNA_snn_res.0.4, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.4) %>%
    mutate(., RNA_snn_res.0.4_cluster_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.4_cluster_purity) %>% distinct() %>% 
    ggplot(., aes(x= reorder(RNA_snn_res.0.4, -RNA_snn_res.0.4_cluster_purity),  y = RNA_snn_res.0.4_cluster_purity))+
    geom_bar(stat = "identity") + plot_sizing +
    ylab("Endo Purity per cluster") + xlab("cluster") 
```

## DotPlot

Use some known marker genes for celltypes to double check scGate annotation.

```{r}
dotplot.level.01 <- c(
  "cdh6",
  "prox1a",
  "stab2",
  "lyve1b",
  "mrc1a",
  "flt1",
  "dll4",
  "esm1",
  "hlx1",
  "hand2",
  "fn1a",
  "gata2b",
  "runx1",
  "alas2",
  "blvrb",
  "spi1b",
  "mpx",
  "neurod4",
  "nova2",
  "cdh1",
  "epcam",
  "myod1",
  "myog",
  "pmela",
  "dct",
  "lum",
  "col1a1a",
  "krt4",
  "cyt1",
  "cryba1b",
  "crybb1",
  "ednrba",
  "pnp4a",
  "gch2",
  "aox5")
```

Make dotplot function:

```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(group)  +plot_sizing_seurat + rotate_x
}

```

Make dotplots for all clusters

```{r, fig.align='center', fig.height=5, fig.width=7}

DimPlot(merged, group.by = "RNA_snn_res.0.4", label = T)
```

```{r, fig.align='center', fig.height=5, fig.width=10}
run_dotplot(merged, dotplot.level.01, group = "RNA_snn_res.0.4")

```
Make dotplot for clusters that have mid-range purity.

```{r}
run_dotplot_mid_purity <- function(seurat, genes, limits){
  clusters <- seurat@meta.data %>% 
  mutate(.,RNA_snn_res.0.4 = as.factor(RNA_snn_res.0.4)) %>% 
  group_by(., RNA_snn_res.0.4, is_endothlial,.drop = F) %>% 
  summarise(., count = n()) %>%
    group_by(RNA_snn_res.0.4) %>%
    mutate(., RNA_snn_res.0.4_cluster_purity = count[is_endothlial == "Pure"]/sum(count)) %>% 
    summarise(RNA_snn_res.0.4_cluster_purity) %>% distinct() %>% 
    filter(., RNA_snn_res.0.4_cluster_purity > limits[1] & RNA_snn_res.0.4_cluster_purity < limits[2]) %>% 
    pull(RNA_snn_res.0.4) %>% unique()
  
  seurat <- SetIdent(seurat, value = "RNA_snn_res.0.4")
  DotPlot(seurat, features = genes, group.by = "RNA_snn_res.0.4", idents = clusters)  +plot_sizing_seurat + rotate_x
}

```

```{r, fig.align='center', fig.height=4, fig.width=10}
run_dotplot_mid_purity(merged, limits = c(0.2, 0.95), genes = dotplot.level.01)
```

They all express other markers as well. 


## Export

```{r}

file_name <- here('output/data/Datsets/merged_wildtypes/merged_wildtypes_Level_01_annotated.qs2')
qs_save(merged, file_name)

```




