---
title: "D10025 yap1 dataset: Make Level 03 (LEC/VEC only) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 03 object for yap1 dataset. Alternative to ./pre_processing_D10025_yap1_dataset_makeObject_Level03.Rmd, but using SCT and regressing out cell cycle instead

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_D10025_yap1_dataset_makeObject_Level02_annotation.Rmd

```{r}
input_paths <- c(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_SCT_kept_doublets_annotated.qs2'),
                 here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_SCT_annotated.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: LEC/VEC clusters only

Subset Level 01 to only include LEC and VEC subsets. Annotation was done in .pre_processing_D10025_yap1_dataset_makeObject_Level02_annotation.Rmd

```{r}
celltype_to_keep <- c("LEC", "VEC")

run_subset_lecvec <- function(seurat, keep = celltype_to_keep){
  cells_to_keep <- seurat@meta.data %>% 
  filter(., L2_celltype %in% celltype_to_keep) %>% 
  rownames_to_column("Barcode") %>% 
  pull(Barcode) 
  
  seurat <- subset(seurat, cells = cells_to_keep)
  return(seurat)
}
```


```{r}
list_objects$D10025_yap1_dataset_Level_03_kept_doublets <- run_subset_lecvec(list_objects$D10025_yap1_dataset_Level_02_SCT_kept_doublets_annotated)
  
list_objects$D10025_yap1_dataset_Level_03 <- run_subset_lecvec(list_objects$D10025_yap1_dataset_Level_02_SCT_annotated)
```

Clean up list, remove Level 02

```{r}
list_objects$D10025_yap1_dataset_Level_02_SCT_kept_doublets_annotated <- list_objects$D10025_yap1_dataset_Level_02_SCT_annotated <- NULL
gc()
```

Replace name in object

```{r}
list_objects$D10025_yap1_dataset_Level_03_kept_doublets@misc$name <- "Level_03_kept_doublets"
list_objects$D10025_yap1_dataset_Level_03@misc$name <- "Level_03"

```


## Re-run single-cell workflow

Not using sctransform, regressing out %mito. Same function we used for Level 01, 02

```{r}
options(future.globals.maxSize = 30000 * 1024^2)
run_workflow <- function(seurat){
  seurat <- SCTransform(seurat, vars.to.regress = c("percent.mt", "tricycle_position")) %>%  #using glmGamPoi 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1))
  
  #add normalisation to RNA as well
  DefaultAssay(object = seurat) <- "RNA"
  seurat <- NormalizeData(seurat) %>%   #should deal with basic library size differences
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt","tricycle_position"))
  
  return(seurat)
}
```


```{r, message=FALSE}
list_objects <- lapply(list_objects, run_workflow)
```


## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F))
```

## UMAP overviews

Make UMAPs for features, ncounts, clusters, genotype, cell cycle, Level 2 celltypes

```{r}
plot_umap_dataset <- function(seurat){
  colours <- Polychrome::dark.colors(5); names(colours) <- c("AEC", "VEC", "LEC", 
                                                             "cycling_EC", "Endocardium")
  
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[5]] <- DimPlot(seurat,  group.by = c("tricycle_stage"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[6]] <- DimPlot(seurat,  group.by = c("L2_celltype"), shuffle = T,cols =colours) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=30}
plot_list <- lapply(list_objects, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F)) +  plot_layout(nrow = 2)
```

Look at predicted doublets for when we kept them:

```{r, fig.align='center', fig.height=5, fig.width=5}
DimPlot(object = list_objects$D10025_yap1_dataset_Level_03_kept_doublets, group.by = "scDblFinder.class")
```

Dont really cluster together.


## Clustree + Composition


```{r, fig.align='center', fig.height=12, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  clustree(seurat, prefix = c("SCT_snn_res.")) + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2)
```

Maybe res 0.4? 0.5 seems to split again. 

Stacked barplot for different resolutions capturing genotype split.

```{r}
run_barplot_genotype <- function(seurat){
  seurat@meta.data %>% 
    select(., Genotype, starts_with("SCT")) %>% 
    pivot_longer(., cols = starts_with("SCT"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("SCT_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, Genotype) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= Genotype)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =2) + 
    strip_fix +
    ggtitle(seurat@misc$name)
}
```


Make UMAPs for res 0.4, 0.5, 0.6

```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- DimPlot(seurat, group.by = "SCT_snn_res.0.4", shuffle = T, label = T) + ggtitle(sprintf("res 0.4, %s", seurat@misc$name)) 
    plot_list[[2]] <- DimPlot(seurat, group.by = "SCT_snn_res.0.5", shuffle = T, label = T) + ggtitle(sprintf("res 0.5, %s", seurat@misc$name)) 
    plot_list[[3]] <- DimPlot(seurat, group.by = "SCT_snn_res.0.6", shuffle = T, label = T) + ggtitle(sprintf("res 0.6, %s", seurat@misc$name)) 
  return(plot_list)

})

wrap_plots(unlist(plot_list, recursive = F), ncol = 3)
```


## scGate: Find major subpopulations

As scGate was no longer useful on Level 02, we're not going to use it on Level 03. We can add it later if we think its necessary.

## Feature UMAPs

```{r}
level_03_dotplot <- c(
  "stab2",  #cvp
  "stab1", #cvp
  "ifi30",#cvp
  "mafba",#cvp
  "cxcl12a", #mapk signaling
  "esm1", #migrating
  "hlx1", #migrating
  "mki67", #cycling
  "pcna", #cycling
  "nop53", #lec? 
  "ddx21", #lec?
  "tbx1", #lec
  "prox1a", #LEC
  "cdh6", #LEC, hippo
  "fn1a", #hippo
  "meox1"
)
```

This can sometimes be helpful.

```{r}
run_feature_umap <- function(seurat, genes){
  FeaturePlot(seurat, features = genes, ncol = 4) 
}

```

```{r, fig.align='center', fig.height=40, fig.width=20}
plot_list <- lapply(list_objects, function(seurat) run_feature_umap(seurat, level_03_dotplot))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

You can see one of the VEC cluster starting to express prox1a, they might just not be proper LECs yet? We could increase the resolution until we get it to split off, but that would give us a very high resolution...

## Dotplot for genes


Make dotplot function:


```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(sprintf("%s, %s",group, seurat@misc$name))  +plot_sizing_seurat + rotate_x
}

```

### Various resolutions{.tabset}

Make for res 0.4, 0.5, 0.6

```{r, fig.align='center', fig.height=8, fig.width=10, results='asis'}
for(res in c("SCT_snn_res.0.4", "SCT_snn_res.0.5","SCT_snn_res.0.6")){
  
  cat(paste0("#### ",res, "\n"))  #make tabs
  
  plot_list <- lapply(list_objects, function(seurat) run_dotplot(seurat, level_03_dotplot, group = res))
  
  wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")

  cat("\n\n") #make tabs
  
}


```



## Add annotations

Can't commit yet.

## Export 

Export Level 03. 

```{r}

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_03_SCT.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_03, file_name)

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_03_SCT_kept_doublets.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_03_kept_doublets , file_name)

```



