---
title: "D10051 meox1 dataset: Make Level 03 (LEC/VEC only) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 03 object for meox1 dataset. 

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_D10051_meox1_dataset_makeObject_Level02_annotation.Rmd

```{r}
input_paths <- c(here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_02_kept_doublets_annotated.qs2'),
                  here('output/data/Datsets/D10051_meox1_dataset/D10051_meox1_dataset_Level_02_annotated.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_keeplegend <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 16)) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: LEC/VEC clusters only

Subset Level 01 to only include LEC and VEC subsets. Annotation was done in pre_processing_D10051_meox1_dataset_makeObject_Level02_annotation.Rmd

```{r}
celltype_to_keep <- c("LEC", "VEC", "pre_muLEC")

run_subset_lecvec <- function(seurat, keep = celltype_to_keep){
  cells_to_keep <- seurat@meta.data %>% 
  filter(., L2_celltype %in% celltype_to_keep) %>% 
  rownames_to_column("Barcode") %>% 
  pull(Barcode) 
  
  seurat <- subset(seurat, cells = cells_to_keep)
  return(seurat)
}
```


```{r}
list_objects$D10051_meox1_dataset_Level_03_kept_doublets <- run_subset_lecvec(list_objects$D10051_meox1_dataset_Level_02_kept_doublets_annotated)
  
list_objects$D10051_meox1_dataset_Level_03 <- run_subset_lecvec(list_objects$D10051_meox1_dataset_Level_02_annotated)
```

Clean up list, remove Level 02

```{r}
list_objects$D10051_meox1_dataset_Level_02_kept_doublets_annotated <- list_objects$D10051_meox1_dataset_Level_02_annotated <- NULL
gc()
```

Replace name in object

```{r}
list_objects$D10051_meox1_dataset_Level_03_kept_doublets@misc$name <- "Level_03_kept_doublets"
list_objects$D10051_meox1_dataset_Level_03@misc$name <- "Level_03"

```


## Re-run single-cell workflow

Not using sctransform, regressing out %mito. Same function we used for Level 01, 02

```{r}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #this is technically unneccessary, but also won't hurt
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1)) #this will overwrite old clustering (that's okay)
}
```


```{r}
list_objects <- lapply(list_objects, run_workflow)
```


## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F))
```

Correlation lib size with PC3.

## UMAP overviews

Make UMAPs for features, ncounts, clusters, genotype, cell cycle, Level 2 celltypes

```{r}
plot_umap_dataset <- function(seurat){
  colours <- Polychrome::dark.colors() %>% unname()
  
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[5]] <- DimPlot(seurat,  group.by = c("tricycle_stage"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[6]] <- DimPlot(seurat,  group.by = c("L2_celltype"), shuffle = T,cols =colours) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=30}
plot_list <- lapply(list_objects, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F)) +  plot_layout(nrow = 2)
```

## Low quality clusters


```{r, fig.align='center', fig.height=8, fig.width=17}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nCount_RNA), x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[2]] <- seurat@meta.data %>% 
  ggplot(., aes(y = log10(nFeature_RNA), x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  plot_list[[3]] <- seurat@meta.data %>% 
  ggplot(., aes(y = percent.mt, x = seurat_clusters, fill=seurat_clusters)) +
  geom_boxplot() +
  plot_sizing + theme(legend.position = "none")
  
  return(plot_list) 
})

wrap_plots(unlist(plot_list, recursive = F)) + plot_layout(axis_titles = "collect") 
```

Not sure if they're low quality or just have a smaller library... from the combined wildtype analysis we see that that celltype doesn't necessarily have a lower library size. The low quality clusters also don't seem doublet related as they appear in both settings. 

## Low quality clusters - do they have markers?

Only running for one object. 

```{r}
markers_low <- FindAllMarkers(object = list_objects$D10051_meox1_dataset_Level_03, 
                              group.by = "seurat_clusters")


markers_low %>% filter(., cluster == 1 & p_val_adj < 0.05)
```

Make scatterplot with % expressed, colour by avglog2FC

```{r, fig.align='center', fig.height=10, fig.width=12}
names <-markers_low %>% 
  filter(., p_val_adj < 0.05) %>% 
  group_by(cluster) %>% 
  summarise(., n_genes = n()) %>% 
  mutate(., name = sprintf("cluster %s\n#DEG=%i", as.character(cluster),n_genes))


markers_low %>% 
  full_join(., names) %>% 
  ggplot(., aes(x=pct.1, y=pct.2, colour=avg_log2FC)) +
  geom_point() +
  facet_wrap(~name, scale="free") +
  plot_sizing_keeplegend+strip_fix +
  scale_colour_gradient2(low="green4", high = "purple4", mid = "grey80")


```


Clusters 1,5, and 7 show a different pattern of DEGs, with the vast majority of DEGs being downregulated. All of this does make it look like these cells are probably low quality and should be removed...

Not continuing, re-run. Keep this for documentation though. 


