---
title: "D10025 yap1 dataset: Paper annotations for all Levels"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

# Introduction

Formal annotation of all Levels using marker genes and scGate scores (for Level 01 only)


## Load libraries

```{r}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
 })

```


## Load gene sets

```{r}
source(here("code/general_scripts/yap1_meox1_important_genes.R"))
```


## Load all objects

Level 01 object contains scGate information. This is the scGate model used:

```{r, eval=FALSE}
endothelial_gating_model <- gating_model(name = "endothelial", signature = c("kdrl+", 
                                                                    "col1a1a-",
                                                                    "epcam-",
                                                                    "gata2b-", 
                                                                    "myod1-"))
```


```{r}

level_01 <- qs_read(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_01_annotated.qs2'))
level_02 <- qs_read(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_annotated.qs2')) #ignore annotations in here and replace
level_03 <- qs_read(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_03.qs2'))

```


## Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))

```


# Annotate Level 01

Resolution 0.5


## get cell numbers per cluster

```{r}
level_01$RNA_snn_res.0.5 %>% table()
```

## Endothelial cell purity

```{r}
level_01@meta.data %>% 
  group_by(RNA_snn_res.0.5,RNA_snn_res.0.5_cluster_purity) %>% 
  summarise()
```




## dotplot

```{r}
DotPlot(object = level_01, features = level_01_annotation_genes, group.by = "RNA_snn_res.0.5") + rotate_x
```

## Add to object

celltype

```{r}
level_01$L1_celltype <- case_when(
  level_01$RNA_snn_res.0.5 %in% c(0,1,2,4,9,11) ~ "Endothelial",
  level_01$RNA_snn_res.0.5 %in% c(8, 16) ~ "Endocardium",
  level_01$RNA_snn_res.0.5 %in% c(3, 13) ~ "Fibroblast",
  level_01$RNA_snn_res.0.5 %in% c(5,10,14, 17,20) ~ "Epithelial",
  level_01$RNA_snn_res.0.5 %in% c(6,7) ~ "Neuronal",
  level_01$RNA_snn_res.0.5 == 12 ~ "Hematopoietic",
  level_01$RNA_snn_res.0.5 == 15 ~ "Chromatophore",
  level_01$RNA_snn_res.0.5 == 18 ~ "Eye_lens",
  level_01$RNA_snn_res.0.5 == 19 ~ "Epithelial_Fibroblast",
  level_01$RNA_snn_res.0.5 == 21 ~ "Immune",
  level_01$RNA_snn_res.0.5 == 22 ~ "Erythrocyte"
)

```

cluster_id

```{r}
names <- c(
'Endothelial_01',
'Endothelial_02',
'Endothelial_03',
'Fibroblast_01',
'Endothelial_04',
'Epithelial_01',
'Neuronal_01',
'Neuronal_02',
'Endocardium_01',
'Endothelial_05',
'Epithelial_02',
'Endothelial_06',
'Hematopoietic_01',
'Fibroblast_02',
'Epithelial_03',
'Chromatophore_01',
'Endocardium_02',
'Epithelial_04',
'Eye_lens_01',
'Epithelial_Fibroblast_01',
'Epithelial_05',
'Immune_01',
'Erythrocyte_01'
)


level_01        <- SetIdent(level_01, value="RNA_snn_res.0.5")
names(names)        <- levels(level_01)
level_01      <- RenameIdents(level_01, names)
level_01      <- AddMetaData(
  object=level_01,
  metadata=level_01@active.ident,
  col.name="L1_cluster_id")

```


## export

```{r}
filename <- here('output/data/Datsets/paper/D10025_yap1_dataset_Level_01_annotated.qs2')
qs_save(object = level_01, file = filename)
```


# Annotate Level 02

Resolution 0.3

## get cell numbers per cluster

```{r}
level_02$RNA_snn_res.0.3 %>% table()
```


## dotplot

```{r}
DotPlot(object = level_02, features = level_02_annotation_genes, group.by = "RNA_snn_res.0.3") + rotate_x
```

## Add to object

celltype

```{r}
level_02$L2_celltype <- case_when(
  level_02$RNA_snn_res.0.3 %in% c(0, 5) ~ "AEC",
  level_02$RNA_snn_res.0.3 %in% c(1,6,8) ~ "VEC",
  level_02$RNA_snn_res.0.3 %in% c(7,10) ~ "cycling_AEC",
  level_02$RNA_snn_res.0.3 == 2 ~ "cycling_EC",
  level_02$RNA_snn_res.0.3 == 4 ~ "migratory_AEC",
  level_02$RNA_snn_res.0.3 == 9 ~ "LEC",
  level_02$RNA_snn_res.0.3 == 3 ~ "Endocardium"
)

```

cluster_id

```{r}
names <- c(
'AEC_01',
'VEC_01',
'cycling_EC_01',
'Endocardium_01',
'migratory_AEC_01',
'AEC_02',
'VEC_02',
'cycling_AEC_01',
'VEC_03',
'LEC_01',
'cycling_AEC_02'
)


level_02       <- SetIdent(level_02, value="RNA_snn_res.0.3")
names(names)        <- levels(level_02)
level_02      <- RenameIdents(level_02, names)
level_02      <- AddMetaData(
  object=level_02,
  metadata=level_02@active.ident,
  col.name="L2_cluster_id")

```



## clean up meta

```{r}
level_02@meta.data <- level_02@meta.data %>% 
  select(-c(starts_with("is.pure"), ends_with("UCell"),"scGate_multi"))
```

## export


```{r}
filename <- here('output/data/Datsets/paper/D10025_yap1_dataset_Level_02_annotated.qs2')
qs_save(object = level_02, file = filename)
```

# Annotate Level 03

Resolution 0.2

## get cell numbers per cluster

```{r}
level_03$RNA_snn_res.0.2 %>% table()
```


## dotplot

```{r}
DotPlot(object = level_03, features = level_03_annotation_genes, group.by = "RNA_snn_res.0.2") + rotate_x
```

## Add to object

celltype

```{r}
names <- c(
'preLEC',
'msVEC',
'hmsVEC',
'cvpVEC',
'LEC',
'IL_VEC'
)


level_03      <- SetIdent(level_03, value="RNA_snn_res.0.2")
names(names)        <- levels(level_03)
level_03      <- RenameIdents(level_03, names)
level_03      <- AddMetaData(
  object=level_03,
  metadata=level_03@active.ident,
  col.name="L3_celltype")

```

cluster_id

```{r}
names <- c(
'preLEC_01',
'msVEC_01',
'hmsVEC_01',
'cvpVEC_01',
'LEC_01',
'IL_VEC_01'
)


level_03      <- SetIdent(level_03, value="RNA_snn_res.0.2")
names(names)        <- levels(level_03)
level_03      <- RenameIdents(level_03, names)
level_03      <- AddMetaData(
  object=level_03,
  metadata=level_03@active.ident,
  col.name="L3_cluster_id")

```



## clean up meta

```{r}
level_03@meta.data <- level_03@meta.data %>% 
  select(-c(starts_with("is.pure"), ends_with("UCell"),"scGate_multi", "L2_celltype", "L2_seurat_celltype"))
```

## export


```{r}
filename <- here('output/data/Datsets/paper/D10025_yap1_dataset_Level_03_annotated.qs2')
qs_save(object = level_03, file = filename)
```

