---
title: "D10025 yap1 dataset: Make Level 03 (LEC/VEC only) object"
author: "Michelle Meier"
date: "`r format(Sys.time(), '%d. %B %Y')`"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

Making Level 03 object for yap1 dataset. 

### Set-up

Load libraries

```{r, warning=FALSE}
suppressPackageStartupMessages({
  library(tidyverse)
  library(here)
  library(Seurat)
  library(scDblFinder)
  library(ggpubr)
  library(clustree)
  library(patchwork)
  library(scater)
  library(scran)
  library(qs2)
  library(scGate)
})
```


### Load data

Made here: ./pre_processing_D10025_yap1_dataset_makeObject_Level02_annotation.Rmd

```{r}
input_paths <- c(here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_kept_doublets_annotated.qs2'),
                 here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_02_annotated.qs2'))
list_objects <- lapply(input_paths, qs_read)

names(list_objects) <- gsub("\\.qs2", "", basename(input_paths))
```

### Plotting themes

```{r}
plot_sizing <- theme_bw()+ theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14),
        legend.title = element_blank()) 

plot_sizing_seurat <- theme(axis.title = element_text(size = 16),
        axis.text = element_text(size = 14, colour = "black"),
        plot.title = element_text(size = 18),
        legend.text = element_text(size = 14))

rotate_x <- theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

strip_fix <- theme(strip.background = element_rect(fill = "white"), 
                   strip.text = element_text(size = 14))
```


## Subset datasets: LEC/VEC clusters only

Subset Level 01 to only include LEC and VEC subsets. Annotation was done in .pre_processing_D10025_yap1_dataset_makeObject_Level02_annotation.Rmd

```{r}
celltype_to_keep <- c("LEC", "VEC")

run_subset_lecvec <- function(seurat, keep = celltype_to_keep){
  cells_to_keep <- seurat@meta.data %>% 
  filter(., L2_celltype %in% celltype_to_keep) %>% 
  rownames_to_column("Barcode") %>% 
  pull(Barcode) 
  
  seurat <- subset(seurat, cells = cells_to_keep)
  return(seurat)
}
```


```{r}
list_objects$D10025_yap1_dataset_Level_03_kept_doublets <- run_subset_lecvec(list_objects$D10025_yap1_dataset_Level_02_kept_doublets_annotated)
  
list_objects$D10025_yap1_dataset_Level_03 <- run_subset_lecvec(list_objects$D10025_yap1_dataset_Level_02_annotated)
```

Clean up list, remove Level 02

```{r}
list_objects$D10025_yap1_dataset_Level_02_kept_doublets_annotated <- list_objects$D10025_yap1_dataset_Level_02_annotated <- NULL
gc()
```

Replace name in object

```{r}
list_objects$D10025_yap1_dataset_Level_03_kept_doublets@misc$name <- "Level_03_kept_doublets"
list_objects$D10025_yap1_dataset_Level_03@misc$name <- "Level_03"

```


## Re-run single-cell workflow

Not using sctransform, regressing out %mito. Same function we used for Level 01, 02

```{r}
run_workflow <- function(seurat){
  seurat <- NormalizeData(seurat) %>%   #this is technically unneccessary, but also won't hurt
    FindVariableFeatures() %>% 
    ScaleData(., vars.to.regress = c("percent.mt")) %>% 
    RunPCA() %>% 
    RunUMAP(., dims = 1:30) %>% #defaults
    FindNeighbors(., dims = 1:30) %>% 
    FindClusters(., resolution = seq(0.1, 1, 0.1)) #this will overwrite old clustering (that's okay)
}
```


```{r}
list_objects <- lapply(list_objects, run_workflow)
```


## PCA correlations

```{r}
run_pca_correlations <- function(seurat){
  plot_list <- list()

  meta <- seurat@meta.data %>% 
    mutate(., log10ncount = log10(nCount_RNA),
              log10nfeat = log10(nFeature_RNA)) %>% 
    select(., log10ncount, log10nfeat) %>% 
    rownames_to_column("Barcode") %>% 
    full_join(., seurat@reductions$pca@cell.embeddings %>% as.data.frame() %>% rownames_to_column("Barcode") %>% select(., Barcode, PC_1, PC_2,PC_3))
  
  #ncount
  plot_list[[1]] <- ggscatter(meta, x = "PC_1", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC1 vs library size, %s", seurat@project.name))
  
  #nfeat
  plot_list[[2]] <- ggscatter(meta, x = "PC_2", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC2 vs library size, %s", seurat@project.name))
  
  #mito
  plot_list[[3]] <- ggscatter(meta, x = "PC_3", y = "log10ncount",
          add = "reg.line",                                 
          conf.int = TRUE,                                
          add.params = list(color = "red4",
                            fill = "lightgray")) +
    plot_sizing + ylab("log10(nCount)") + xlab("PC1") +
    stat_cor(method = "pearson", size = 5, colour = "red4") +
    ggtitle(sprintf("PC3 vs library size, %s", seurat@project.name))
  
  return(plot_list)
}
```


```{r, fig.align='center', fig.height=10, fig.width=15}
plot_list <- lapply(list_objects, run_pca_correlations)

wrap_plots(unlist(plot_list, recursive = F))
```

Correlation lib size with PC3.

## UMAP overviews

Make UMAPs for features, ncounts, clusters, genotype, cell cycle, Level 2 celltypes

```{r}
plot_umap_dataset <- function(seurat){
  colours <- Polychrome::dark.colors(5); names(colours) <- c("AEC", "VEC", "LEC", 
                                                             "cycling_EC", "Endocardium")
  
  plot_list <- list()
  plot_list[[1]] <- FeaturePlot(seurat,  features = c("nCount_RNA")) + ggtitle(seurat@project.name)
  plot_list[[2]] <- FeaturePlot(seurat, features = c("nFeature_RNA")) + ggtitle(seurat@project.name)
  plot_list[[3]] <- DimPlot(seurat,  group.by = c("seurat_clusters")) + ggtitle(seurat@project.name)
  plot_list[[4]] <- DimPlot(seurat,  group.by = c("Genotype"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[5]] <- DimPlot(seurat,  group.by = c("tricycle_stage"), shuffle = T) + ggtitle(seurat@project.name)
  plot_list[[6]] <- DimPlot(seurat,  group.by = c("L2_celltype"), shuffle = T,cols =colours) + ggtitle(seurat@project.name)
  return(plot_list)
  
}
```

Top with, bottom without doublets. 

```{r, fig.align='center', fig.height=10, fig.width=30}
plot_list <- lapply(list_objects, plot_umap_dataset)

wrap_plots(unlist(plot_list, recursive = F)) +  plot_layout(nrow = 2)
```

## Clustree + Composition


```{r, fig.align='center', fig.height=12, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  clustree(seurat) + ggtitle(seurat@misc$name)
})

wrap_plots(plot_list, ncol = 2)
```

Reassignments in kept doublets at low resolutions. With removed doublets its a lot cleaner. 

Stacked barplot for different resolutions capturing genotype split.

```{r}
run_barplot_genotype <- function(seurat){
  seurat@meta.data %>% 
    select(., Genotype, starts_with("RNA")) %>% 
    pivot_longer(., cols = starts_with("RNA"), values_to = "cluster", names_to = "resolution") %>% 
    mutate(., resolution = gsub("RNA_snn_", "", resolution)) %>% 
    group_by(resolution, cluster, Genotype) %>% 
    summarise(., count = n()) %>% 
    ggplot(., aes(x= cluster, y = count, fill= Genotype)) +
    geom_bar(position = "fill", stat = "identity") +
    plot_sizing +
    facet_wrap(~resolution, scales = "free_x", nrow =2) + 
    strip_fix +
    ggtitle(seurat@misc$name)
}
```


```{r, fig.align='center', fig.height=20, fig.width=30}

plot_list <- lapply(list_objects, run_barplot_genotype)

wrap_plots(plot_list, ncol = 1)
```


Make UMAPs for res 0.2, 0.3, 0.4

```{r, fig.align='center', fig.height=10, fig.width=20}
plot_list <- lapply(list_objects, function(seurat){
  plot_list <- list()
  plot_list[[1]] <- DimPlot(seurat, group.by = "RNA_snn_res.0.2", shuffle = T, label = T) + ggtitle(sprintf("res 0.2, %s", seurat@misc$name)) 
  plot_list[[2]] <- DimPlot(seurat, group.by = "RNA_snn_res.0.3", shuffle = T, label = T) + ggtitle(sprintf("res 0.3, %s", seurat@misc$name)) 
  plot_list[[3]] <- DimPlot(seurat, group.by = "RNA_snn_res.0.4", shuffle = T, label = T) + ggtitle(sprintf("res 0.4, %s", seurat@misc$name)) 
    plot_list[[4]] <- DimPlot(seurat, group.by = "RNA_snn_res.0.5", shuffle = T, label = T) + ggtitle(sprintf("res 0.5, %s", seurat@misc$name)) 
  return(plot_list)

})

wrap_plots(unlist(plot_list, recursive = F), ncol = 4)
```

Look at predicted doublets for when we kept them:

```{r, fig.align='center', fig.height=5, fig.width=5}
DimPlot(object = list_objects$D10025_yap1_dataset_Level_03_kept_doublets, group.by = "scDblFinder.class")
```

They clearly cluster together. 


## scGate: Find major subpopulations

As scGate was no longer useful on Level 02, we're not going to use it on Level 03. We can add it later if we think its necessary.

## Dotplot for genes

```{r}
level_03_dotplot <- c(
  "stab2",  #cvp
  "stab1", #cvp
  "ifi30",#cvp
  "mafba",#cvp
  "cxcl12a", #mapk signaling
  "esm1", #migrating
  "hlx1", #migrating
  "mki67", #cycling
  "pcna", #cycling
  "nop53", #lec? 
  "ddx21", #lec?
  "tbx1", #lec
  "prox1a", #LEC
  "cdh6", #LEC, hippo
  "fn1a", #hippo
  "meox1"
)
```

Make dotplot function:


```{r}
run_dotplot <- function(seurat, genes, group){
  DotPlot(seurat, features = genes, group.by = group) + ggtitle(sprintf("%s, %s",group, seurat@misc$name))  +plot_sizing_seurat + rotate_x
}

```

### Various resolutions{.tabset}

Make for res 0.2-0.5

```{r, fig.align='center', fig.height=8, fig.width=10, results='asis'}
lapply(c("RNA_snn_res.0.2", "RNA_snn_res.0.3","RNA_snn_res.0.4", "RNA_snn_res.0.5"), function(res){
  
  cat(paste0("#### ",res, "\n"))  #make tabs
  
  plot_list <- lapply(list_objects, function(seurat) run_dotplot(seurat, level_03_dotplot, group = res))
  wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
  
  cat("\n\n") #make tabs
})


```

## Feature UMAPs

This can sometimes be helpful.

```{r}
run_feature_umap <- function(seurat, genes){
  FeaturePlot(seurat, features = genes, ncol = 4) 
}

```

```{r, fig.align='center', fig.height=40, fig.width=20}
plot_list <- lapply(list_objects, function(seurat) run_feature_umap(seurat, level_03_dotplot))

wrap_plots(plot_list, ncol = 1) + plot_layout(guides = "collect", axis_titles = "collect")
```

You can see one of the VEC cluster starting to express prox1a, they might just not be proper LECs yet? We could increase the resolution until we get it to split off, but that would give us a very high resolution...

## Add annotations

Can't commit yet.

## Export 

Export Level 03. 

```{r}

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_03.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_03, file_name)

file_name <- here('output/data/Datsets/D10025_yap1_dataset/D10025_yap1_dataset_Level_03_kept_doublets.qs2')
qs_save(list_objects$D10025_yap1_dataset_Level_03_kept_doublets , file_name)

```



